### 2025-02-28 paper/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Space for WHAM: a multi-region, multi-stock generalization of the Woods Hole Assessment Model with an application to black sea bass},
  pdfauthor={Timothy J. Miller1; Kierstin Curti; Alexander C. Hansell},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{url}
\usepackage{setspace}
%\singlespacing
%\onehalfspacing
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage[belowskip=0pt,aboveskip=0pt]{caption}
\usepackage{relsize}
\usepackage{float}
% \usepackage[section]{placeins}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{amsmath,rotating}
\usepackage[scanall]{psfrag}
\usepackage{bm}
\usepackage{caption,graphics}
\usepackage{graphicx}
%\usepackage{natbib}
%\usepackage[nottoc]{tocbibind}
%\usepackage{indentfirst}
\usepackage{sectsty}
\usepackage{color}
\usepackage{fancyhdr}
\usepackage{xspace}
\usepackage{textcomp}
\usepackage{upgreek}

\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\newcommand{\noop}[1]{}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Space for WHAM: a multi-region, multi-stock generalization of the
Woods Hole Assessment Model with an application to black sea bass}
\author{Timothy J. Miller\textsuperscript{1} \and Kierstin
Curti \and Alexander C. Hansell}
\date{28 February, 2025}

\begin{document}
\maketitle

\(^1\)\href{mailto:timothy.j.miller@noaa.gov}{\nolinkurl{timothy.j.miller@noaa.gov}},
Northeast Fisheries Science Center, National Marine Fisheries Service,
166 Water Street, Woods Hole, MA 02543, USA\\

\pagebreak

\hypertarget{main-message}{%
\subsection*{Main Message}\label{main-message}}
\addcontentsline{toc}{subsection}{Main Message}

Describes the multi-region, multi-stock generalization of WHAM and its
usage to evaluate evidence of alternative hypotheses about temperature
effects on recruitment and age 1 natural mortality of black sea bass
stock components.

\hypertarget{abstract}{%
\subsection*{Abstract}\label{abstract}}
\addcontentsline{toc}{subsection}{Abstract}

The Woods Hole Assessment Model is a general state-space age-structured
assessment model that is used to assess and manage many stocks in the
Northeast US. We first describe an extension of the model allowing any
number of stocks (or stock components) and and regions with movement
among regions as well as seasonal variation in stock and fleet dynamics.
Movement rates can be a be functions of time- and age-varying random
effects and environmental covariates. We then illustrate the model by
applying it to data for the northern and southern components of the
Northeast US black sea bass stock and evaluate alternative hypotheses of
bottom temperature and time-varying random effects on recruitment and
natural mortality. We show strong evidence for temperature effects on
recruitment, primarily for the northern stock component, and no evidence
for including random effects or temperature effects on age 1 natural
mortality.

\pagebreak

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

A state-space age-structured stock assessment models can be used to
estiamte time and age-varying population attributes as random effects
using maximum marginal likelihood or Bayesian fitting procedures
(Nielsen and Berg 2014; Cadigan 2016; Miller et al. 2016b). This
estimation approach is considered an essential feature of gold-standard
assessment models that we use in tactical management of commercially
important fish stocks (Punt et al. 2020). The State-space Assessment
Model (SAM, Nielsen and Berg 2014) continues to be developed and remains
widely used within the International Council for the Exploration of the
Sea (ICES) to assess European fish stocks. Various state-space models
are being used to manage cod and plaice stocks in the waters of Eastern
Canada (Perreault et al. 2020; Varkey et al. 2022) and to the south, the
Woods Hole Assessment Model (WHAM, Miller and Stock 2020; Stock and
Miller 2021) is now used to assess many fish stocks in the Northwest
Atlantic Ocean (NEFSC 2022a, 2022b; NEFSC 2024).

WHAM is an R package developed and maintained at NOAA's Northeast
Fisheries Science Center {[}\url{https://timjmiller.github.io/wham}{]}.
WHAM can be configured to fit a wide range of age-structured models from
traditional statistical catch-at-age models without any random effects
to models with several time and age varying process errors and effects
of environmental covariates on various population parameters. Like SAM,
WHAM models are built using the Template Model Builder package (TMB,
Kristensen et al. 2016) which provides a computationally efficient means
of fitting an extremely wide class of models with random effects. WHAM
has undergone active development since its creation and includes random
effects options for recruitment, inter-annual transitions of numbers at
age (hereafter referred to as ``apparent survival''), fishery and index
selectivity, natural mortality, and catchability.

However, WHAM has only allowed one stock (component) and a single region
and without seasonal changes in stock dynamics. Using such models for
stocks that have subcomponents with varying seasonal movement can
provide incorrect inferences and poor management advice (Ying et al.
2011; Cao et al. 2014; Bosley et al. 2022). Furthermore, the ability to
account for spatial structure and model multiple stocks are also
important features of leading-edge assessment modeling frameworks (Punt
et al. 2020). We describe here the implementation of these features and
other extensions since Stock and Miller (2021) in WHAM version 2.0. Many
of these new configuration options can also be useful when modeling a
single stock and region.

We developed this extension of WHAM in concert with a research track
assessment process for black sea bass (\textit{Centropristis striata})
(NEFSC 2023). Previous black sea bass assessments assumed two stock
components (North and South), each assessed with a separate statistical
catch at age assessment. The stock was split into two regions to account
for spatial dynamics and improve model diagnostics (ASMFC 2016).
Additionally, black sea bass have complex spatial dynamics and make
seasonal migrations, moving inshore in the spring and offshore in the
fall (Moser and Shepherd 2009). Black sea bass distribution has been
linked to warming waters on the northwest shelf (Bell et al. 2015) and
it is hypothesized that the distribution and productivity of the black
sea bass stock is especially susceptible to long-term changes in
temperature causing increases in abundance further north (Hare 2016).
Thus, a focus of the research track assessment was to develop an
assessment platform that could capture multiple stock components,
complex seasonal movement and explore environmental drivers of
population dynamics (NEFSC 2023).To illustrate the usage of WHAM 2.0 we
apply it to two stock components of black sea bass off the coast of the
NEUS using most of the model assumptions that were accepted during the
peer-review process. However, here we evaluate evidence for alternative
hypotheses of temporal variation and effects specifically of bottom
temperature on recruitment and natural mortality of age 1 individuals.

When modeling multiple stock components occurring in different regions
further dimensionality is added to the class of possible equilbrium
reference points (e.g., Kapur et al. 2021). Short term projections are
also an important part of the management process and the separation of
observation and process errors in state-space models allows uncertainty
in projected attributes such as recruitment and spawning stock biomass
(SSB) without external simulation exercises. Finally, the inclusion of
explicit effects of environmental covariates in the model also requires
some assumptions on how they are treated in projections and reference
points. We demonstrate some of options of WHAM version 2.0 to provide
biological reference points and short-term projections.

\hypertarget{methods}{%
\section*{Methods}\label{methods}}
\addcontentsline{toc}{section}{Methods}

\hypertarget{wham-description}{%
\subsection*{WHAM description}\label{wham-description}}
\addcontentsline{toc}{subsection}{WHAM description}

Many of the options and equations of WHAM version 2.0 are the same as
those in Stock and Miller (2021), so we will only describe extensions
and differences that have occurred since their first description of
WHAM. The new version of WHAM can model multiple stocks each with their
own movement, harvest, and natural mortality. Much of the description
below is for a specific stock \(s\), but, for simplicity, this subscript
is implicit except when necessary.

\hypertarget{the-probability-transition-matrix}{%
\subsubsection*{The probability transition
matrix}\label{the-probability-transition-matrix}}
\addcontentsline{toc}{subsubsection}{The probability transition matrix}

Because individuals may be alive in one of several regions or harvested
in one of several fleets, it is helpful to consider these as distinct
categories or states and treat the number of individuals occurring in
each category over time as a multi-state model. Approaches to modeling
transitions among these states may treat time discretely (e.g., Arnason
1972; Schwarz et al. 1993) or continuously (e.g., Hearn et al. 1987;
Commenges 1999; Andersen and Keiding 2002). We can define a probability
transition matrix (PTM) that describes the probability of individuals
occurring in different states at the end of a time interval \(\delta\)
(columns), conditional on being in each of those state at beginning of
the interval (rows). For fish populations, these states would be defined
as being alive in a particular region or deceased due to fishing from a
particular fleet or natural mortality. The time interval \(i\) with
duration \(\delta_i\) would be a season and the PTMs would be uniquely
defined for each stock by season \(i\), year \(y\), and age \(a\) on
January 1. Each row and column of the PTM correspond to one the states:
alive in region \(r\), dead in fleet \(f\), or dead from natural causes.
The probabilities in each row sum to unity (must be in one of the
states) and assume an individual is in the state that corresponds to
that row at the beginning of the interval. Given \(n_R\) regions and
\(n_F\) fleets, the square PTM (\(n_R + n_F + 1\) rows and columns) as a
function of sub-matrices is \begin{equation}\label{eq:ptm}
  \mathbf{P}_{y,a,i} = \begin{bmatrix}
    \mathbf{O}_{y,a,i} & \mathbf{H}_{y,a,i} & \mathbf{D}_{y,a,i} \\
    0 & \mathbf{I}_{H} & 0\\
    0 & 0 & 1
  \end{bmatrix}
\end{equation} where \begin{equation*}
  \mathbf{O}_{y,a,i} = 
  \begin{bmatrix}
    O_{y,a,i}(1,1) & \cdots & O_{y,a,i}(1,n_R) \\
    \vdots & \ddots & \vdots \\
    O_{y,a,i}(n_R,1) & \cdots & O_{y,a,i}(n_R,n_R)
  \end{bmatrix}
\end{equation*} is the \(n_R \times n_R\) matrix defining survival and
occurring in each region at the end of the interval, \begin{equation*} 
  \mathbf{H}_{y,a,i} = 
  \begin{bmatrix}
    H_{y,a,i}(1,1) & \cdots & H_{y,a,i}(1,n_F) \\
    \vdots & \ddots & \vdots \\
    H_{y,a,i}(n_R,1) & \cdots & H_{y,a,i}(n_R,n_F)
  \end{bmatrix}
\end{equation*} is the \(n_R \times n_F\) matrix defining probabilities
of being captured in each fleet during the interval, and
\(\mathbf{D}_{y,a,i}\) is the \(n_R x 1\) matrix of probabilities of
dying due to natural mortality during the interval. We have the identity
matrix \(\mathbf{I}_{H}\) for the states of capture by each fleet and a
1 for the natural mortality state because the probabilities of being in
one of the mortality states given starting the interval in that state is
unity (no zombies allowed).

WHAM uses these PTMs to model abundance proportions in each state rather
than true probabilities where numbers in each state would be multinomial
distributed as in a model for tagging data where fates of individual
tagged fish are assumed independent. The PTMs determine the expected
numbers 1) in each state on January 1 of year \(t+1\) at age \(a+1\)
given the abundances at age \(a\) on January 1 of year \(t\), 2)
captured over the year in each fleet, 3) available to each index, and 4)
alive at the time and in the region where spawning occurs.

\hypertarget{single-region-ptms}{%
\subsubsection*{Single region PTMs}\label{single-region-ptms}}
\addcontentsline{toc}{subsubsection}{Single region PTMs}

When there is only one region, \begin{equation}\label{eq:ptm_1_region}
\mathbf{P}_{y,a,i} = 
  \begin{bmatrix}
     S_{y,a,i} & \mathbf{H}_{y,a,i} & D_{y,a,i} \\
     0 & \mathbf{I}_{H} & 0\\
     0 & 0 & 1
  \end{bmatrix}
\end{equation} where \(S_{y,a,i} = e^{-Z_{y,a,i}\delta_i}\),
\(\mathbf{H}_{y,a,i}\) is a 1 x \(n_F\) matrix with elements for each
fleet \(f\):
\(\frac{F_{y,a,i,f}}{Z_{y,a,i}}\left(1 - e^{-Z_{y,a,i}\delta_i}\right)\),
\(D_{y,a,i} = \frac{M_{y,a}}{Z_{y,a,i}}\left(1 - e^{-Z_{y,a,i}\delta_i}\right)\),
and \(Z_{y,a,i} = M_{y,a} + \sum^{n_F}_{f=1} F_{y,a,i,f}\) is the total
mortality rate.

\hypertarget{multi-region-ptms}{%
\subsubsection*{Multi-region PTMs}\label{multi-region-ptms}}
\addcontentsline{toc}{subsubsection}{Multi-region PTMs}

When there is more than one region, WHAM can model survival and movement
as processes occurring sequentially or simultaneously. The sequential
assumption is used widely in spatially explicit model (e.g., Stock
Synthesis, Methot and Wetzel 2013). Under the sequential assumption,
survival and death occur over the interval and movement among regions
occurs instantly at either the beginning or the end of the interval.
WHAM is configured to have movement occur after survival:
\begin{equation*}
  \mathbf{O}_{y,a,i} = \mathbf{S}_{y,a,i}\boldsymbol{\mu}_{y,a,i}
\end{equation*} where \(\mathbf{S}_{y,a,i}\) is a \(n_R \times n_R\)
diagonal matrix of proportions surviving in each region (given they
start in that region) \begin{equation*}
\mathbf{S}_{y,a,i} = 
  \begin{bmatrix}
    e^{-Z_{y,a,i,1}\delta_i} & 0 & \cdots & 0 \\
    0 & e^{-Z_{y,a,i,2}\delta_i} & \cdots & 0 \\
    \vdots & \vdots & \ddots & \vdots \\
    0 & \cdots & 0 & e^{-Z_{y,a,i,n_R}\delta_i}
  \end{bmatrix}
\end{equation*} and \(\boldsymbol{\mu}_{y,a,i}\) is a \(n_R \times n_R\)
matrix of probabilities of moving from one region to another or staying
in the region they occurred at the beginning of the interval:
\begin{equation*}
\boldsymbol{\mu}_{y,a,i} = 
  \begin{bmatrix}
    1-\sum_{r' \neq 1} \mu_{1\rightarrow r',y,a,i} & \mu_{1\rightarrow 2,y,a,i} & \cdots & \mu_{1\rightarrow R,y,a,i} \\
    \mu_{2\rightarrow 1,y,a,i} & 1-\sum_{r' \neq 2} \mu_{2\rightarrow r',y,a,i} & \cdots & \mu_{2\rightarrow R,y,a,i} \\
    \vdots & \vdots & \ddots & \vdots \\
    \mu_{R\rightarrow 1,y,a,i} & \cdots & \mu_{R\rightarrow R-1,y,a,i} & 1-\sum_{r' \neq R} \mu_{R\rightarrow r',y,a,i}
  \end{bmatrix}
\end{equation*}

WHAM assumes each fleet \(f\) can harvest in only one region (\(r_f\))
during specified seasons. So, for each fleet \(f\), row \(r_f\) and
column \(f\) of \(\mathbf{H}_{y,a,i}\) will be
\(F_{y,a,i,f}\left(1 - e^{-Z_{y,a,i,r}\delta_i}\right)/Z_{y,a,i,r}\)
when fleet \(f\) is harvesting during the interval \(\delta_i\) and all
other elements will be zero. Row \(r\) of the single-column matrix
\(\mathbf{D}_{y,a,i}\) is
\(M_{y,a,r}\left(1 - e^{-Z_{y,a,i,r}\delta_i}\right)/Z_{y,a,i,r}\)

When survival and movement are assumed to occur simultaneously, all
movement and mortality parameters are instantaneous rates. We obtain the
probability transition matrix over an interval \(\delta_i\) by
exponentiating the instantaneous rate matrix (Miller and Andersen 2008)
\begin{equation*}
\mathbf{P}_{y,a,i} = e^{\mathbf{A}_{y,a,i}\delta_i}
\end{equation*} The instantaneous rate matrix takes rates of movement
between regions and the mortality rates for each fleet and region. Along
the diagonal is the negative of the sum of the other rates (the hazard)
so each row sums to zero. For two regions and one fleet operating in
each region: \begin{equation*}
 \mathbf{A}_{y,a,i} = \begin{bmatrix}
 a_{y,a,i,1} & \mu_{1\rightarrow 2,y,a,i} & F_{y,a,i,1} & 0 & M_{y,a,1} \\
 \mu_{2\rightarrow 1,y,a,i} &  a_{y,a,i,2} & 0 & F_{y,a,i,2} & M_{y,a,2} \\
 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0
 \end{bmatrix}.
\end{equation*} where
\(a_{y,a,i,r} = -(\mu_{r\rightarrow r',y,a,i} + F_{y,a,i,r_f} + M_{y,a,r})\).
When there is one region, \(n_f\) fleets, and \(\delta_i = 1\),
exponentiating the instantaneous rate matrix results in the PTM defined
in Eq. \ref{eq:ptm_1_region}.

\hypertarget{seasonality}{%
\subsubsection*{Seasonality}\label{seasonality}}
\addcontentsline{toc}{subsubsection}{Seasonality}

Seasonality can be configured to accommodate characteristics of
spawning, movement, and fleet-specific behavior. The annual time step
can be divided into \(K\) seasons and the interval size \(\delta_i\) for
each season \(i\) does not need to be equal to any other seasonal
interval. Under the Markov assumption, the PTM of surviving and moving
and dying over \(K\) intervals \(\delta_1,\ldots, \delta_K\) (i.e., the
entire year) is just the product of the PTMs for each interval:
\[ \mathbf{P}_{y,a}(\delta_1,\ldots,\delta_K) = \prod^K_{i=1}\mathbf{P}_{y,a,i}(\delta_i).\]

For a stock spawning at some fraction of the year \(0<t_s<1\) in
interval \(\delta_j\), the fraction of time in season \(j\) is
\[\delta_{s,j} = t_s-\sum^{j-1}_{i=0}\delta_i
\] and the PTM defining the proportions in each state at time \(t_s\)
for age \(a\) is \begin{equation}\label{eq:ptm_spawn}
\mathbf{P}_{y,a}\left(\delta_1,\ldots,\delta_{j-1}, \delta_{s,j}\right) = \mathbf{P}_{y,a}\left(t_s\right) =  \left[\prod^{j-1}_{i=1}\mathbf{P}_{y,a,i}(\delta_i)\right]\mathbf{P}_{y,a,j}(\delta_{s,j}).
\end{equation} Similarly, for an index \(m\) occurring at fraction of
the year \(t_m\) in interval \(\delta_j\) the proportions in each state
at the time of the observation is \begin{equation} \label{eq:ptm_index} 
\mathbf{P}_{y,a}\left(\delta_1,\ldots,\delta_{j-1}, \delta_{m,j}\right) = \mathbf{P}_{y,a}\left(t_m\right) =   \left[\prod^{j-1}_{i=1}\mathbf{P}_{y,a,i}(\delta_i)\right]\mathbf{P}_{y,a,j}(\delta_{m,j}).
\end{equation}

\hypertarget{numbers-at-age}{%
\subsubsection*{Numbers at age}\label{numbers-at-age}}
\addcontentsline{toc}{subsubsection}{Numbers at age}

When there are \(n_R\) regions and \(n_F\) fleets, the vector of
abundance in each state at age \(a>1\) on January 1 is
\(\mathbf{N}_{y,a} = (\mathbf{N}_{O,y,a}', \mathbf{0}')'\) where
\(\mathbf{N}_{O,y,a} = (N_{y,a,1}, \ldots, N_{y,a,n_R})'\) is the number
in the states corresponding to being alive in each region and
\(\mathbf{0}\) is a vector (\(n_F+1\)) for the numbers captured in each
fleet and dead from natural mortality because no age \(a\) fish have
died yet on January 1.

Each stock \(s\) is assumed to spawn and recruit in one region \(r_s\).
So for age \(a=1\), \(\mathbf{N}_{O,y,1}\) is 0 except for row
\(r = r_s\). Options for configuring recruitment (\(N_{y,1,r_s}\)) for
each stock are the same as previous versions of WHAM (Miller and Stock
2020). If recruitment is assumed to be a function of SSB, it is only the
spawning population in region \(r_s\) at the time of spawning that
constitutes the SSB in the stock-recruit function. However, models can
configure spawning individuals to occur in other regions at the time of
spawning. Aside from treating recruitment as a random walk, the general
model for annual recruitment as random effects is \begin{equation*}
\log\left(N_{y,1,r_s}\right)|\text{SSB}_{y-1,r_s} =  f\left(\text{SSB}_{y-1,r_s}\right) + \varepsilon_{y,1,r_s}
\end{equation*} where\\
\[\text{SSB}_{y,r_s} =  \sum^A_{a=1}  w_{y,a} \text{mat}_{y,a} \mathbf{O}_{y,a,r_s}(t_s)' \mathbf{N}_{O,y,a}\]
and \(w_{y,a}\) is the mean weight at age of spawning individuals,
\(\text{mat}_{y,a}\) is the maturity at age, and
\(\mathbf{O}_{y,a,r_s}(t_s)\) are the probabilities of surviving and
occurring in region \(r_s\) at time \(t_s\) given being alive in each
region at the start of the year (the \(r_s\) column of the upper-left
submatrix of Eq. \ref{eq:ptm_spawn}).

As in previous versions of WHAM, the transitions in numbers at age from
one year to another after recruitment can be treated deterministically
or as functions of random effects. The predicted numbers at age in year
\(y\) at age \(a\) for a given stock are vector analogs
(\(\mathbf{N}_{O,y,a}\)) of the equations for numbers at age in the
standard WHAM model (Stock and Miller 2021). For ages
\(a = 2,\ldots, A-1\), where \(A\) is the plus group, the expected
number alive in each region at the beginning of the following year and
next age class age can be obtained from the first \(n_R\) elements of
the vector \[\mathbf{P}_{y-1,a-1}' \mathbf{N}_{y-1,a-1}.\] The numbers
alive in each region can also be modeled more simply using the
sub-matrix \(\mathbf{O}_{y,a}\). The general model for the transitions
in abundance at age is \begin{equation*}
\log\left(\mathbf{N}_{O,y,a}\right)|\mathbf{N}_{O,y-1,a-1} =  \log\left(\mathbf{O}_{y-1,a-1}' \mathbf{N}_{O,y-1,a-1}\right) + \boldsymbol{\varepsilon}_{y,a}
\end{equation*} for ages \(a = 2,\ldots, A-1\), and for the plus group
\begin{equation*}
\log\left(\mathbf{N}_{O,y,A}\right)|\mathbf{N}_{O,y-1,a-1},\mathbf{N}_{O,y-1,A} = \log\left(\mathbf{O}_{y-1,A-1}' \mathbf{N}_{O,y-1,A-1} + \mathbf{O}_{y-1,A}' \mathbf{N}_{O,y-1,A}\right) + \boldsymbol{\varepsilon}_{y,A}
\end{equation*} where \(\boldsymbol{\varepsilon}_{y,a}\) is the vector
of region-specific errors for a given stock, which are independent
across stocks and regions, but the same correlation structures as
previous versions are possible across ages and years for a given stock
and region. When there is autocorrelation with age, WHAM now assumes
this applies only to ages \(a>1\) by default so that recruitment random
effects are independent of those for for the annual transitions of older
age classes. So the general covariance structure for a given stock at
ages \(a>1\) in region \(r\) is that of a two-dimensional first-order
autoregressive (2DAR1) process \begin{equation*}
  Cov\left(\varepsilon_{y,a,r},\varepsilon_{y',a',r}\right) =   \frac{\rho_{N,\text{age},r}^{|a-a'|}\rho_{N,\text{year},r}^{|y-y'|}\sigma_{N,a,r}\sigma_{N,a',r}}{\left(1 -  \rho_{N,\text{age},r}^2\right)\left(1 - \rho_{N,\text{year},r}^2\right)} 
\end{equation*} and that for age 1 is just AR1 across years
\begin{equation*}
  Cov\left(\varepsilon_{y,1},\varepsilon_{y',1}\right) =   \frac{\rho_{N,1,\text{year}}^{|y-y'|}\sigma^2_{N,1}}{1 - \rho_{N,1,\text{year}}^2}.
\end{equation*} When the transitions in abundance at age are treated
deterministically, \(\boldsymbol{\varepsilon}_{y,a} = 0\). Since
recruitment for a given stock currently only occurs in one region
\(r_s\) there is only a single vector of time-varying recruitment random
effects for each stock.

\hypertarget{initial-numbers-at-age}{%
\subsubsection*{Initial numbers at age}\label{initial-numbers-at-age}}
\addcontentsline{toc}{subsubsection}{Initial numbers at age}

Initial numbers at age for each stock and region can be treated as
age-specific fixed effects or with an equilibrium assumption as in
previous versions of WHAM. For the equilibrium option there are two
parameters for each stock: the stock-specific fully-selected fishing
mortality rate \(\log \widetilde{F}\) and the recruitment in year 1
\(\log N_{1,1,r_s}\). A stock-specific equilibrium fishing mortality at
age by fleet \({\tilde F}_{a,f}\) is the product of \(\widetilde{F}\)
and the selectivity across fleets in the first year \begin{equation*}
  \text{sel}_{1,a,f} = \frac{F_{1,a,f}}{\max_a \sum_{f=1}^{n_F} F_{1,a,f}}.
\end{equation*} We use \(\widetilde{F}_{a,f}\) to define an equilibrium
abundance per recruit by region at age \(a\) conditional on recruiting
to each region \begin{equation}\label{eq:npraa}
 \widetilde{\mathbf{O}}_{a} = \left\{
 \begin{array}{ll}
\prod^{a-1}_{j=0}\mathbf{O}_{j}  & 1\leq a<A\\
\left[\prod^{a-1}_{j=0}\mathbf{O}_{j}\right] \left(\mathbf{I} - \mathbf{O}_{A}\right)^{-1} & a = A
 \end{array}
\right.
\end{equation} where \(\mathbf{O}_{j}\) is the equilibrium probability
of surviving age \(a\) and occurring in each region and
\(\mathbf{O}_{0} = \mathbf{I}\). Natural mortality and movement rates in
the first year of the model are also used in Eq. \ref{eq:npraa}. For the
plus group \(a=A\), \(\left(\mathbf{I} - \mathbf{O}_{A}\right)^{-1}\) is
a ``fundamental matrix'\,' derived using the matrix version of the
geometric series (Kemeny and Snell 1960). Recall that recruitment for
stock \(s\) only occur in region \(r_s\) so, the equilibrium initial
numbers at age \(a\) by region are
\[\mathbf{N}_{O,1,a} = \widetilde{\mathbf{O}}_{a}' \mathbf{N}_{O,1,1}.\]

The initial abundances at age can also be treated as independent or AR1
random effects. Defining the vector of initial abundance at age in
region \(r\) as \(\mathbf{N}_{O,1,r}\), the general model is
\[\log \mathbf{N}_{O,1,r} = \theta_{N_1,r} + \boldsymbol{\varepsilon}_{N_1,r}\]
where
\[  Cov\left(\varepsilon_{N_1,a,r},\varepsilon_{N_1,a',r}\right) = \frac{\rho_{N_1,r}^{|a-a'|}\sigma^2_{N_1,r}} {\left(1-\rho_{N_1,r}^2\right)}.\]

\hypertarget{parametizing-movement}{%
\subsubsection*{Parametizing movement}\label{parametizing-movement}}
\addcontentsline{toc}{subsubsection}{Parametizing movement}

For each season, there are at most \(n_R-1\) parameters determining
movement among regions given starting the season in region \(r\) in
either the sequential or simultaneous configurations. Movement
parameters are estimated on a transformed scale via a link function
\(g(\cdot)\). If survival and movement occur simultaneously, the
parameters are estimated with a log link function and if they are
separable, an additive logit link function (like a multinomial
regression) is used. On the transformed scale, the general model for the
movement parameter from region \(r\) to \(r'\) in season \(i\) and year
\(y\) for individuals of age \(a\) is a linear function of an intercept
or mean parameter (\(\theta_{r\rightarrow r',i}\)) and both random and
environmental effects: \begin{equation*}
  g(\mu_{r\rightarrow r',y,a,i}) = \theta_{r\rightarrow r',i} + \varepsilon_{r\rightarrow r',y,a,i} + \sum^{n_E}_{k=1} \beta_{r \rightarrow r',a,i,k} E_{k,y}.
\end{equation*} The random effects
\(\varepsilon_{r\rightarrow r',y,a,i}\) are season-, and
region-to-region-specific and modeled most generally as 2DAR1 random
effects with age and(or) year where the covariance is \begin{equation*}
  Cov\left(\varepsilon_{r\rightarrow r',y,a,i},\varepsilon_{r\rightarrow r',y',a',i}\right) =   \frac{\rho_{r\rightarrow r',\text{age},i}^{|a-a'|}\rho_{r\rightarrow r',\text{year},i}^{|y-y'|}\sigma^2_{r\rightarrow r',i}}{\left(1 -  \rho_{r\rightarrow r',\text{age},i}^2\right)\left(1 - \rho_{r\rightarrow r',\text{year},i}^2\right)}
\end{equation*} similar to how WHAM models variation in survival,
natural mortality, and selectivity. Effects of covariate \(E_k\) can be
age-, season-, and region-to-region-specific
\(\beta_{r\rightarrow r',a,i,k}\) and the same orthogonal polynomial
options in the previous versions of WHAM for effects on recruitment and
natural mortality are available.

There is currently no likelihood component for tagging data. Therefore,
movement parameters would usually either need to be fixed or assumed to
have some prior distribution, possibly based on external parameter
estimates. WHAM includes prior distributions for the season and
region-to-region specific (mean) movement parameters and treats them as
random effects with the mean defined by the initial value of the fixed
effect counterpart and standard deviation \begin{equation*}
  \gamma_{r\rightarrow r',i} \sim \text{N}\left(\theta_{r\rightarrow r',i}, \sigma^2_{r\rightarrow r',i}\right).
  \end{equation*} When priors are used, the movement is defined instead
as \begin{equation*}
  g(\mu_{r\rightarrow r',y,a,i}) = \gamma_{r\rightarrow r',i} + \varepsilon_{r\rightarrow r',y,a,i} + \sum^{n_E}_{k=1} \beta_{r\rightarrow r',a,i,k} E_{k,y} 
  \end{equation*}

\hypertarget{natural-mortality}{%
\subsubsection*{Natural mortality}\label{natural-mortality}}
\addcontentsline{toc}{subsubsection}{Natural mortality}

When treated as constant parameters, natural mortality rates may be
stock-, region-, and age-specific. When random effects are used, the
same 2DAR1 structure with age and year as described by Stock and Miller
(2021) can be configured for a given stock and region and any
environmental covariate effects can be stock-, region-, and
age-specific. So the general model for natural mortality is
\begin{equation*}
  \log M_{y,a,r} = \theta_{M,r} + \varepsilon_{M,r,y,a} + \sum^{n_E}_{k=1} \beta_{M,r,a,k} E_{k,y}.
\end{equation*} The general covariance structure for random effects are
modeled most generally as 2DAR1 random effect with age and(or) year
where the covariance is \begin{equation*}
  Cov\left(\varepsilon_{M,y,a,r},\varepsilon_{M,y',a',r}\right) =   \frac{\rho_{M,\text{age},r}^{|a-a'|}\rho_{M,\text{year},r}^{|y-y'|}\sigma^2_{M,r}}{\left(1 -  \rho_{M,\text{age},r}^2\right)\left(1 - \rho_{M,\text{year},r}^2\right)}.
\end{equation*}

\hypertarget{catch-observations}{%
\subsubsection*{Catch observations}\label{catch-observations}}
\addcontentsline{toc}{subsubsection}{Catch observations}

The log-normal distributional assumption for aggregate catch
observations is the same as Stock and Miller (2021), but the predicted
catch is now a function of catch from each stock starting the year in
each region. For a given stock and age, the numbers captured in each
fleet over the year are
\[\widehat{\mathbf{N}}_{H,s,y,a} = \mathbf{H}_{s,y,a}' \mathbf{N}_{O,s,y,a}\]
The predicted numbers caught be each fleet across stocks is
\[\widehat{\mathbf{N}}_{H,y,a} = \sum^{n_S}_{s=1} \widehat{\mathbf{N}}_{H,s,y,a}\]
and the predicted aggregate catch at age \(a\) is
\[\widehat{\mathbf{C}}_{y,a} = \text{diag}\left(\mathbf{c}_{y,a}\right) \widehat{\mathbf{N}}_{H,y,a}\]
where \(\mathbf{c}_{y,a}\) is the vector of mean individual weight at
age \(a\) for each fleet and the aggregate catch by fleet is
\[\widehat{\mathbf{C}}_y = \sum^{A}_{a=1} \widehat{\mathbf{C}}_{y,a}\]
The log-aggregate catch observations for fleet \(f\) are normally
distributed
\[ \log C_{y,f} \sim \text{N}\left(\log \widehat {C}_{y,f}, \sigma^2_{y,f}\right).\]

The predicted numbers caught for each fleet \(f\) (row \(f\) of
\(\widehat{\mathbf{N}}_{H,y,a}\)) are used to make predicted age
composition observations as described by Stock and Miller (2021). Since
then, three additional likelihood options for age composition
observations have been added: a logistic-normal with AR(1) correlation
structure (Francis 2014), the alternative Dirichlet-multinomial
parameterization described by Thorson et al. (2017), and the
multivariate Tweedie (Thorson et al. 2023).

\hypertarget{index-observations}{%
\subsubsection*{Index observations}\label{index-observations}}
\addcontentsline{toc}{subsubsection}{Index observations}

For index \(m\) occurring in region \(r_m\) at fraction of the year
\(t_m\), the predicted abundance at \(t_m\) in region \(r_m\) is
\[\widehat{N}_{s,y,a,m} = \mathbf{O}_{s,y,a,r_m}(t_m)' \mathbf{N}_{O,s,y,a}\]
where \(\mathbf{O}_{s,y,a,r_m}(t_m)\), the \(r_m\) column of the
upper-left submatrix of Eq. \ref{eq:ptm_index}, are the probabilities of
surviving and occurring in region \(r_m\) at time \(t_m\) given being
alive in each region at the start of the year. The predicted index at
age is
\[\widehat{I}_{m,y,a} = q_{m,y} \text{sel}_{m,y,a}w_{m,y,a}\sum^{n_S}_{s = 1}\widehat{N}_{s,y,a,m}\]
where \(q_{m,y}\) is the catchability of the index in year \(y\),
\(\text{sel}_{m,y,a}\) is the selectivity and \(w_{m,y,a}\) is the
average weight of individuals at age \(a\) if the index is quantified in
biomass and \(w_{m,y,a} = 1\) if the index is quantified in numbers.
Predicted age composition observations are functions of
\(\widehat{I}_{m,y,a}\) as described by Stock and Miller (2021) and the
likelihood options are the same as those for catch explained above.

Catchability of the index can also be treated as logit functions of
normal random effects and(or) environmental covariate effects
\[log \frac{q_{m,y}-l_m}{u_m-q_{m,y}} = \theta_{q,m} + \varepsilon_{q,m,y}  + \sum^{n_E}_{k=1} \beta_{q,m,k} E_{k,y}\]
where \(\theta_{q,m}\) is an intercept or mean parameter, and \(u_{m}\)
and \(l_{m}\) are the upper and lower bounds of catchability for index
\(m\) (defaults are 0 and 1000). The general covariance structure for
the annual random effects is AR1
\[Cov\left(\varepsilon_{q,m,y},\varepsilon_{q,m,y'}\right) =   \frac{\rho_{q,m}^{|y-y'|}\sigma^2_{q,m}}{1 - \rho_{q,m}^2}.\]

\hypertarget{reference-points}{%
\subsubsection*{Reference points}\label{reference-points}}
\addcontentsline{toc}{subsubsection}{Reference points}

Currently a single fishing mortality reference point \(\widetilde F\) is
estimated across stocks and regions and the fleet- and age-specific
reference points is
\(\widetilde F_{f,a} = \widetilde F \text{sel}_{f,a}\). Selectivity is
determined across fleets and ages as in the equilibrium assumption for
initial abundance at age except here it is a function of average F at
age over a user-defined set of years \begin{equation}\label{eq:spr_sel}
  \text{sel}_{a,f} = \frac{\overline F_{a,f}}{\max_a \sum^{n_F}_{f=1}{\overline F}_{a,f}}
\end{equation}

The equilibrium spawning stock biomass per recruit for a given stock in
region \(r_s\) is defined as \begin{equation}\label{eq:ssbpr}
 \upphi(\widetilde{F}) = \sum^{A}_{a=1} \widetilde{\mathbf{O}}_{a,r_s,\cdot} \mathbf{O}_{a,\cdot,r_s}(t_s) w_{a} m_{a}
\end{equation} where \(w_{a}\) and \(m_{a}\) are the mean individual
weight and probability of maturity at age \(a\),
\(\widetilde{\mathbf{O}}_{a}\) are as described in Eq. \ref{eq:npraa},
and \(\mathbf{O}_{a}(t_s)\) is the \(n_R \times n_R\) upper-left
sub-matrix of eq. \ref{eq:ptm_spawn} with the probabilities of surviving
and occurring in each region \(r'\) at age \(a+t_s\) given starting in
region \(r\) at age \(a\). The further subscripts \(r_s,\cdot\) and
\(\cdot,r_s\) indicate row or column \(r_s\), respectively. Using these
rows and columns is required because of the assumption that spawning and
recruitment only occur in region \(r_s\).

The equilibrium spawning biomass per recruit (eq. \ref{eq:ssbpr}) is
conditional on the region of recruitment \(r_s\). The equilibrium
recruitment depends on the stock dynamics. WHAM currently only allows
complete spawning region fidelity so that a stock only spawns and
recruits in a single region. Therefore, the equilibrium vector of
numbers at age 1 for each state, \(\widetilde {\mathbf{N}}_{1}\), will
only be nonzero in the row corresponding to the spawning region
(\(r_s\)). The matrices of probabilities of surviving and occurring in
each region, \(\widetilde{\mathbf{O}}_{s,a}\) and
\(\mathbf{O}_{s,a}(t_s)\), are functions of the fishing mortality rates
for fleets in each region \(\widetilde F_{f,a}\).

The matrix of equilibrium yield per recruit by region and fleet as a
function of \(\widetilde F\) is calculated as
\begin{equation}\label{eq:ypr}
 \widetilde{\nu}({\widetilde{F}}) = \sum^{A}_{a=1} \widetilde{\mathbf{O}}_{a,r_s,\cdot} \mathbf{H}_{a} \mathbf{c}_{s,a}
\end{equation} where \(\mathbf{c}_a\) is the vector of mean individual
weight at age for each fleet, and \(\mathbf{H}_{s,a}\) is the submatrix
of the probabilities of being captured in each fleet over the interval
from \(a\) to \(a+1\), defined in eq. \ref{eq:ptm}.

As in previous versions of WHAM package, reference points for prevailing
conditions, average all of the inputs to the spawning biomass and yield
per recruit calculations (weight, maturity, natural mortality, and
movement) over a set of user-specified years as shown in eq.
\ref{eq:spr_sel} for selectivity (e.g., last 5 years of the model).

For \(X\%\) SPR-based reference points, we use a Newton method and
iterate \begin{equation}\label{eq:newton}
  \log\widetilde{F}^{(i)} = \log\widetilde{F}^{(i-1)} - \frac{g\left(\log\widetilde{F}^{(i-1)}\right)}{g'\left(\log\widetilde{F}^{(i-1)}\right)}
\end{equation} where \(g(\log F)\) is the difference between the
weighted sums of spawning biomass per recruit at \(F\) and \(X\)\% of
unfished spawning biomass per recruit across stocks:
\begin{equation}\label{eq:newton-obj}
  g(\log F) = \sum^{n_s}_{s=1} \lambda_s\left[\upphi_s\left(F = e^{\log F}\right) - \frac{X}{100}\upphi_s\left(F=0\right)\right].
\end{equation} where \(\upphi_s\left(F=0\right)\) is the equilibrium
unfished spawning biomass per recruit. \(g'(\log F)\) is the derivative
of \(g\) with respect to \(\log F\), and the weights to use for each
stock \(\lambda_s\) can be specified by the user or relative to the
average of recruitment for each stock over the same years the user
defines to calculate ``static'\,' equilibrium spawning biomass and
yield.

When a Beverton-Holt or Ricker stock recruit relationship is assumed, an
analogous Newton method is used to find \(\log F\) that maximizes yield
for MSY-based reference points. which are also a functions of the
equilibrium yield per recruit (Eq. \ref{eq:ypr}) and equilibrium
recruitment. The function \(g(\log F)\) in Eq. \ref{eq:newton} is the
first derivative of the yield curve with respect to \(\log F\).

\hypertarget{projections}{%
\subsubsection*{Projections}\label{projections}}
\addcontentsline{toc}{subsubsection}{Projections}

The projection options are generally the same as those for previous
versions of WHAM (Miller and Stock 2020). When there is movement of any
stocks, the user has the option to project and use any random effects
for time-varying movement or use the average over user specified years,
analogous to how natural mortality can be treated in the projection
period. The projection of any environmental covariates has been revised
to better include error in the estimated latent covariate in any effects
on the population in projection years. Users can also specify a catch or
fully-selected fishing mortality in each projection year and they can be
fleet-specific.

\hypertarget{application-to-black-sea-bass}{%
\section*{Application to black sea
bass}\label{application-to-black-sea-bass}}
\addcontentsline{toc}{section}{Application to black sea bass}

Black sea bass are a temperate reef fish in the western Atlantic Ocean
ranging from the Gulf of Mexico to the Gulf of Maine. Fish north of Cape
Hatteras, NC, are considered to comprise a single stock unit, but
individual populations exhibit spawning site fidelity (Able and Stanton
Hales, Jr. 1997; Fabrizio et al. 2013). Fish in this northern stock unit
perform seasonal migrations out on the continental shelf in the fall and
back to their inshore spawning areas during the spring (Musick and
Mercer 1977). Analyses of tagging studies have demonstrated that the
extent of seasonal migration varies along the coast such that fish from
populations off of southern New England and further north move offshore
and as far south as the coasts of Virgina and North Carolina, whereas
fish in the southern portion of the stock range generally move shorter
distances between inshore and offshore areas of similar latitude (Moser
and Shepherd 2009).

Prior to its 2023 peer-reviewed assessment, the NEUS black sea bass
stock was assessed using the Age-Structured Assessment Program (ASAP)
model (Legault and Restrepo 1999), a single-stock and -region
statistical catch-at-age model that estimates all model parameters as
fixed effects. Northern and southern components of the NEUS black sea
bass stock ascribed to regions divided by the Hudson Canyon were
separately modeled in ASAP (reference map figure). Results from the
separate ASAP models were combined for a unit-stock assessment. The
ASAP-based assessments exhibited strong retrospective patterns (Mohn
1999), and exploring alternative modeling approaches for the northern
and southern stock components has been a high priority for management.

Leading up to the 2023 peer-reviewed assessment, a working group
(hereafter referred to as ``Working Group'') composed of scientists from
federal, state, and academic institutions determined an optimal data and
model configuration for the black sea bass stock using the multi-stock
and multi-region extension of WHAM described above (NEFSC 2023). This
assessment included the spatial features and investigated inclusion of
hypothesized environmental drivers that were prioritized research
recommendations from previous black sea bass assessments. Below we
describe the assumptions and configuration of the assessment model as
determined by the Working Group as well as the alternative assumptions
for recruitment and natural mortality in the models we fit to evaluate
alternative hypotheses of bottom temperature effects on black sea bass.

\hypertarget{basic-structure}{%
\subsection*{Basic structure}\label{basic-structure}}
\addcontentsline{toc}{subsection}{Basic structure}

The first year being modeled for the population is 1989 and the fishery
and index data used in the model span from 1989 to 2021. The north and
south stock components are modeled as separate populations that spawn
and recruit in respective regions. We have observations for each of four
total fishing fleets, where two fishing fleets (recreational and
commercial) operate in each region.

There are 11 ``seasonal'' intervals within each calendar year: five
monthly time intervals from Jan 1 to May 31, a spawning season from June
1 to July 31, and five monthly intervals from August 1 to December 31.
The southern stock component is assumed to never move to the northern
region. For the northern component, a proportion
\(\mu_{\text{N}\rightarrow \text{S}}\) can move to the southern region
each month during the last five months of the year, but no movement is
allows from the south to the north during this period (Figure
\ref{fig:migration-diagram}). During the first four intervals of the
year a proportion \(\mu_{\text{S}\rightarrow \text{N}}\) the northern
component individuals in the south can move back to the north, but no
movement from the north to south is allowed during this period. In the
fifth interval (May), all northern component individuals remaining in
the south are assumed to move back to the north for the subsequent
spawning period. Survival and movement occur sequentially in each
interval and each of the two movement proportions are assumed constant
across intervals, ages, and years.

The two monthly movement matrices are \begin{equation*}
\boldsymbol{\mu}_{1} = 
  \begin{bmatrix}
     1-\mu_{\text{N}\rightarrow \text{S}} & \mu_{\text{N}\rightarrow \text{S}} \\
     0 & 1 \\
  \end{bmatrix}
\end{equation*} for the portion of the year after spawning and
\begin{equation*}
\boldsymbol{\mu}_{2} = 
  \begin{bmatrix}
     1 &  0 \\
     \mu_{\text{S}\rightarrow \text{N}} & 1-\mu_{\text{S}\rightarrow \text{N}} \\
  \end{bmatrix}
\end{equation*} for the portion of the year before spawning. As noted in
the description of the general WHAM model, tagging data are not yet
allowed. However, the Working Group also fit a Stock Synthesis model
(Methot and Wetzel 2013) which included tagging data and provided
estimated movement rate parameters and standard errors that were used to
configure the priors for WHAM (see Supplementary Materials).

With movement as configured, the northern origin fish (ages 2+) can
occur in the southern region on January 1. Estimating initial numbers at
age as separate parameters can be challenging even in single-stock
models, but for black sea bass the available data cannot distinguish the
proportion of northern and southern component fish at each age in the
southern region in the initial year of the model. Therefore, we used the
simplifying equilibrium assumption where there are two parameters
estimated for each stock component: an initial recruitment and an
equilibrium fully-selected \(F\) that determines the abundance at age in
each region for each stock component.

For subsequent years, all models allow AR1 correlation of recruitment
and 2DAR1 correlation for survival for both the northern and southern
stock components. Models with apparent survival random effects will
model the transitions (conditional on both survival and movement) of
abundances at age of northern origin fish in each region, but there
appears to be little ability of the existing data from the southern
region to distinguish apparent survival random effects for each stock
component. All models we considered assume very small variance for the
transitions of northern fish in the southern region, which is
essentially the same as the deterministic transition assumptions of a
statistical catch at age model. Unique variance and correlation
parameters for the recruitment and apparent survival random effects are
estimated for the northern and southern components.

\hypertarget{uncertainty-in-recreational-index-observations}{%
\subsection*{Uncertainty in recreational index
observations}\label{uncertainty-in-recreational-index-observations}}
\addcontentsline{toc}{subsection}{Uncertainty in recreational index
observations}

The estimated coefficients of variation (CVs) provided by the analyses
by the Working Group to generate the recreational catch per angler
(recreational CPA) index ranged between 0.02 and 0.06 which the Working
Group felt did not capture the true observation uncertainty in the index
with regard to its relationship to stock abundance. All models we
considered allowed a scalar multiple of the standard deviation of the
log aggregate index to be estimated for these indices in the northern
and southern regions. However, lack of model convergence was observed in
many fits for self-test simulations. The model estimates are negligibly
affected by estimating these scalars, but estimating rather than fixing
these parameters allowed uncertainty in model output to be more properly
conveyed.

\hypertarget{index-and-catch-age-composition-observations}{%
\subsection*{Index and catch age composition
observations}\label{index-and-catch-age-composition-observations}}
\addcontentsline{toc}{subsection}{Index and catch age composition
observations}

The Working Group investigated many alternative assumptions for the
probability models and selectivity forms for the eight different sets of
age composition observations to reduce residual patterns and
retrospective patterns. These analyses resulted in use of selectivity
random effects for the northern fleet and indices and logistic-normal
likelihoods for six sets of of age composition observations and
Dirichlet-multinomial likelihoods for one index and one fleet in the
northern region (Table \ref{tab:age-comp-sel-table}).

\hypertarget{bottom-temperature-effects}{%
\subsection*{Bottom temperature
effects}\label{bottom-temperature-effects}}
\addcontentsline{toc}{subsection}{Bottom temperature effects}

All models include bottom temperature observations for the northern and
southern regions from 1963 to 2021 and standard errors of observations
ranged between 0.03 and 0.09 degrees Celsius (NEFSC 2023). We retained
the assumption from the peer-reviewed assessment that treated the latent
bottom temperature covariates in each region as AR1 processes.

We fit 14 models with alternative assumptions about the effects of
bottom temperature covariates, ranging from no effects to effects on
both regions for either recruitment or natural mortality at age 1 (Table
\ref{tab:model-desc-table}). These analyses derive from the hypothesis
that bottom temperature affects overwinter survival of fish where the
fish turn from age 0 to age 1 on January 1 (Miller et al. 2016a). This
covariate may be a proxy for temperature prior to January 1 and affect
survival during the end of the pre-recruit phase or natural mortality in
the early part of of the year after becoming age 1. Furthermore, we have
no direct observations of age 1 individuals from surveys until the
spring season each year. Therefore, we fit models with effects of
temperature on recruitment or natural mortality at age 1.

All models treat annual recruitment as AR1 time-varying random effects
with mean parameters for each stock component. However, Miller et al.
(2018) showed how inferences of temperature effects on growth or
maturity parameters can be very different whether the compared models
with and without the effect also include random effects representing
residual temporal variation in parameters. Therefore, we also explored
whether including temporal random effects on age 1 natural mortality
affected inferences on corresponding temperature effects. Initially, we
included random effects on age 1 natural mortality for both the northern
and southern stock components, but estimates of these random effects and
corresponding variance for the southern component converged to 0 so
these were not included in models presented here.

We assume the covariate in year \(y\) affects recruitment in the same
years because the covariate observations are from months January to
March. The fish are technically already 1 year old, but there are no
observations of these individuals until later in the year except
possibly in fishery catches which are accumulated over the whole year.
Expected log-recruitment for a given stock is a linear function of
bottom temperature

\begin{equation}\label{eq:expected-recruitment}
E\left(\log N_{y,1}|x_y\right) = \mu_{R} + \beta_{R} x_y.
\end{equation} Similarly, expected log-natural morality as a function of
bottom temperature is \begin{equation}\label{eq:expected-M1}
E\left(\log M_{y,1}|x_y\right) = \mu_{M,1} + \beta_{M} x_y
\end{equation} Because age 1 fish for the northern component can exist
in both regions after January 1, natural mortality is acting in both
regions for this stock. For models with covariate and(or) annual random
effects for age 1 fish we assume them only in the northern regions for
the northern component. The corresponding random effects are
\begin{equation}\label{eq:Rec-re}
\log N_{y,1} = E\left(\log N_{y,1}|x_y\right) + \varepsilon_{y,1}
\end{equation} and \begin{equation}\label{eq:M-re}
\log M_{y,1} = E\left(\log M_{y,1}|x_y\right) + \varepsilon_{M,y}.
\end{equation} The constant or mean log-natural mortality rate is
assumed to be \(\mu_{M,a} = \log(0.4)\) for all ages as recommended by
the Working Group. Because the bottom temperature anomalies and random
effects are centered at 0, mean log-natural mortality at age 1 over the
time series should be approximately equal to \(\mu_{M,1}\) for models
that include those effects.

\hypertarget{model-fitting-and-diagnostics}{%
\subsection*{Model fitting and
diagnostics}\label{model-fitting-and-diagnostics}}
\addcontentsline{toc}{subsection}{Model fitting and diagnostics}

We used a development version
(\href{https://github.com/timjmiller/wham/tree/fb8b089}{commit fb8b089})
of the WHAM package prior to the release of version 2.0 for all results.
All code for fitting models and generating results can be found at
\href{https://github.com/timjmiller/wham_bsb_paper}{github.com/timjmiller/wham\_bsb\_paper}.

We examined retrospective patterns for all models by fitting
corresponding models where the terminal year is reduced sequentially by
one year (peel) for seven years. Therefore, there are eight fits of each
model with the time series reduced by zero to seven years. We calculated
Mohn's \(\rho\) for SSB, and average \(F\) at ages 6 and 7. Absolute
values of Mohn's \(\rho\) near 0 imply no pattern in estimation of these
quantities as the time-series is sequentially extended. As in Miller et
al. (2016b), we also assessed the consistency of the AIC-based model
selection over retrospective peels to guard against previously noted
changes in perception of covariate effects on recruitment with increased
length of the time series of observations (Myers 1998). This
retrospective examination was also recommended by Brooks (2024).

We performed a jitter analysis of the base model \(M_0\) and the best
fitting model to investigate whether a local mimimum of the negative
log-likelihood surface was obtained by the optimization. We used the
jitter\_wham function in the WHAM package which by default simulates
starting values from a multivariate normal distribution with mean and
covariance defined by the MLEs and hessian-based covariance matrix from
the fitted model. See the Supplementary Materials for more details.

For the best performing model, we also performed simulation self-tests
where new observations were simulated conditional on all estimated fixed
and random effects and the same model configuration was fit to each of
the simulated data sets. We estimated median relative bias of SSB across
these simulations. See the Supplementary Materials for more details.

\hypertarget{reference-points-and-projections}{%
\subsection*{Reference points, and
projections}\label{reference-points-and-projections}}
\addcontentsline{toc}{subsection}{Reference points, and projections}

We produced SPR-based BRP estimates (\(F_{40\%}\) and SSB\(_{40\%}\))
from the best performing model using a variety of options provided in
WHAM. We generated annual BRP estimates based on the annual weight at
age, maturity at age, natural mortality, fleet selectivity, and either
the annual recruitment random effect (Eq. \ref{eq:Rec-re}) or the annual
expected recruitment (Eq. \ref{eq:expected-recruitment}). We also
calculated status as the ratios of annual average F and SSB to the
corresponding reference points. Finally, we calculated ``prevailing''
BRPs using the average weight at age, maturity at age, M, fleet
selectivity from the last 5 years (2017 to 2021) and average recruitment
over years 2000 to 2021, and made Kobe plots of joint status for the
terminal year of the model (2021). All BRPs and status estimates are
generated for each region as well as for the total stock.

We projected the best black sea bass model under three alternative
scenarios for the bottom temperature covariate where 1) the AR1
time-series model continues into the projection years, 2) the average of
the most recent 5 years is projected, and 3) bottom temperature
increases in projection years following a prediction from a simple
linear regression of the estimated bottom temperature anomalies over
time. We specified fishing mortality in projection years to be constant
at the value in the last model year (2021). Average weight, natural
mortality, movement, and maturity at age over the last 5 years are used
in projection years whereas the AR1 processes for all numbers at age
random effects are continued.

\hypertarget{results}{%
\section*{Results}\label{results}}
\addcontentsline{toc}{section}{Results}

We found the best model of bottom temperature covariate effects included
the effect only on recruitment of the northern stock and no random
effects on age 1 natural mortality (\(M_1\), Tables
\ref{tab:diff-aic-table} and \ref{tab:aic-wts-table}). Model \(M_1\)
consistently had the lowest AIC across retrospective peels where the
terminal years of data are removed from the model. The difference in AIC
for the model that also included bottom temperature effects on
recruitment for the southern component (\(M_3\)) suggested some evidence
for this hypothesis as well. There was little evidence for also
including random effects on age 1 natural mortality for the northern
stock component. However, when more than three terminal years of data
were removed, the evidence for adding age 1 natural mortality random
effects for the northern component (\(M_8\)) increased is similar to
adding bottom temperature effects on southern component recruitment
(\(M_3\)).

Comparisons of estimates from retrospective peels did not show any
indication of differences in model performance. There was very little
variation in measures of retrospective patterns for SSB or fishing
mortality across models. Mohn's \(\rho \approx -0.03\) for SSB of both
stock components and Mohn's \(\rho \approx 0.03\) and \(\approx -0.04\)
for average \(F\) at ages 6 and 7 in north and south regions,
respectively (Table \ref{tab:rho-table}).

Estimated effects of bottom temperature on recruitment for either the
northern or southern component were stable over the retrospective peels
and differed negligibly whether effects for each component were
estimated in isolation or together (Table
\ref{tab:beta-sig-peel-table}). The estimates for the effects on each
stock component were essentially equivalent whether the effect was on
one component or both. Estimates of residual variability in northern
component recruitment, as measured by the conditional or marginal
standard deviation of the recruitment random effects, increased slightly
with the number of peeled years. However, the ratio of standard
deviations of models with and without temperature effects was stable
with approximately 20\% reduction in residual standard deviation when
temperature effects were included. The residual variation is reduced
because the expected recruitment (Eq. \ref{eq:expected-recruitment}) is
a function of the covariate (Figure \ref{fig:BT-Ecov-R}).

The posterior estimates of the bottom temperature covariate match the
observations well because of the high precision of the observed
anomalies (Figure \ref{fig:bottom-temperature}). The anomalies for the
north and south regions appear highly correlated presumably because of
the proximity of the two regions. Because the temperature anomalies are
treated as latent variables, other data components in the model can
affect the estimated anomalies when effects on recruitment are included
(e.g., Miller et al. 2018), but in this case the estimates of anomalies
are altered negligibly during the population model years when the effect
on recruitment is estimated (Figure \ref{fig:Ecov-M1-rel-M0}).

The effect of the anomalies on expected recruitment can be better
observed by comparing the time series of the northern and southern
components (Figure \ref{fig:E-R-SSB-F}). The expected recruitment for
the southern component without any covariate effects is constant over
time whereas the effect on expected recruitment for the northern
component induces correlation of the bottom temperature anomalies and
expected recruitment. For the northern stock component, most of the
lowest estimated recruitment random effects occurred prior to 2000 and
several large recruitment estimates occurred after 2010. For the
southern stock, recruitment random effect estimates have been less
variable and similar to the estimated median particularly since 2005.
The estimated stock size of the northern component, as measured by SSB,
increased substantially since 2005, whereas that for the southern
component peaked in 2002. The increase in the northern component is also
reflected in the total size of the combined components. Estimated
average total fishing mortality rates across ages 6 and 7 (combining the
commercial and recreational fleets) were highest in the southern region
prior to 1998 and have declined substantially since then. Estimated
fishing mortality rates in the northern region have been less variable,
but had largest peaks in 1996, 1999, and 2005. The decline in estimates
in the southern component is also reflected in the decline in total
fishing mortality across regions.

The estimated movement from the north to the south for the northern
stock component is much less than anticipated from the prior
distribution that was parameterized based on the companion Stock
Synthesis model fit completed by the Working Group (Figure
\ref{fig:move-prior-posterior}). Conversely, the estimated movement rate
from south to north was more consistent and slightly greater than that
from the companion model. The estimated movement rates varied little
across all fitted models implying that the estimation of movement was
not sensitive to the alternative assumptions we considered. Furthermore,
the very low estimated movement rate from north to south is also
consistent with preliminary fits of the WHAM model without the prior
distribution where the movement rate was estimated at the lower bound of
0 resulting in a lack of convergence. The estimated movement rates imply
that a very small fraction of northern fish are ever in the southern
region.

The 2DAR1 random effects for the selectivity parameters of the northern
recreational fleet indicate a decrease in selectivity at ages 2 to 6
over the time series which corresponds to increases in the size limits
in regulations by states where these catches occurred (Figure
\ref{fig:selectivity-re}). The model also estimated a more modest
decrease in selectivity at ages 2 to 3 for the northern commercial
fleet. The selectivity at the youngest ages for the two fishery
independent indices varied over time but without notable trend.

\hypertarget{reference-points-1}{%
\subsection*{Reference points}\label{reference-points-1}}
\addcontentsline{toc}{subsection}{Reference points}

Using the expected recruitment rather than the random effects as the
multiplier with equilibrium spawning biomass per recruit
(\(\upphi(\widetilde{F})\), eq. \ref{eq:ssbpr}) results in less variable
estimates of annual SSB\(_{40\%}\) (Figure \ref{fig:annual-BRPs}: top
row). Estimates were less variable for the southern stock component than
the northern stock component mainly because the expected recruitment is
constant for the former whereas that for the northern component varies
via the effect of the bottom temperature anomalies. Temporal variation
for the southern stock component is mainly caused by annual variation in
the inputs to the \(\upphi(\widetilde{F})\) calculation and
\(F_{40\%}\), but there is also some effect of the variation in the
proportion of total recruitment for each stock component as used in the
weighting terms to define the 40\% reduction in unfished spawning
biomass per recruit (\(\lambda_s\) in Eq. \ref{eq:newton-obj}). Using
the recruitment random effects instead of expected recruitment creates
more uncertain estimates of SSB\(_{40\%}\) for the southern stock
component whereas there is little difference for the northern component
because there is less difference in the types of recruitment estimates
resulting from the covariate effects (Figure
\ref{fig:annual-SSB40-cvs}).

Ratios of SSB to SSB\(_{40\%}\) should be viewed with caution because
the SSB reference point is being defined using annual recruitments
rather than some form of average recruitment and therefore, ratios
larger than 2.5 which would correspond to unfished population sizes are
possible (Figure \ref{fig:annual-BRPs}: second row). However, for the
southern component, using the constant expected recruitment is
comparable to using median recruitment and the annual variation in the
inputs to \(\upphi(\widetilde{F})\) indicates no years where SSB was
greater than unfished levels throughout the time series. The patterns in
uncertainty of SSB ratios are the same as those for the SSB reference
points (Figure \ref{fig:annual-SSB-status-cvs}).

Differences between annual \(F_{40\%}\) estimates using the alternative
types of recruitment estimates are much less than those for
SSB\(_{40\%}\) because the recruitment estimates are only used in the
weighting of the stock-specific \(\upphi(\widetilde{F})\) to determine
\(F_{40\%}\) (\(\lambda_s\) in Eq. \ref{eq:newton-obj}, Figure
\ref{fig:annual-BRPs}: third row). For similar reasons, estimates of
uncertainty of \(F_{40\%}\) are also similar between the type of
recruitment estimates (Figure \ref{fig:annual-F40-cvs}). Annual
estimates for the northern stock component were lowest in the earliest
years of the time series whereas those for the soutern component are
relatively stable over time.

For a given type of recruitment used to calculate \(F_{40\%}\), the
annual ratios of average \(F\) to the corresponding reference point are
identical for each stock component and in total because of the
selectivity used to define a single F reference point across both
regions is exactly the same as that for the annual estimate of fishing
mortality at age by fleet (Figure \ref{fig:annual-BRPs}: fourth row). As
for annual \(F_{40\%}\), estimates of uncertainty in the ratios are
similar whether expected recruitment or the random effects are used to
calculate the reference points (Figure \ref{fig:annual-F-status-cvs}).

Unlike the annual ratios of \(F\) to \(F_{40\%}\), the \(F_{40\%}\)
under prevailing conditions uses a selectivity defined by the average
\(F\) at age and fleet over the last 5 years, which is different from
that of the selectivity in the terminal year and therefore those ratios
for each region and in total are different (Figure
\ref{fig:annual-BRPs}: bottom row). Although the alternative types of
recruitment estimates result in different points on the Kobe plots for
joint status, the uncertainty in the estimates suggest similar
probabilities of alternative quadrants. However, the usage of the
confidence regions of the ratios to defined probabilities is an
approximation because these ratios are functions of both empirical Bayes
estimates of random effects as well as maximum likelihood estimates of
fixed effects.

\hypertarget{projections-1}{%
\subsection*{Projections}\label{projections-1}}
\addcontentsline{toc}{subsection}{Projections}

When continuing the AR1 model for the bottom temperature anomalies in
the projection years, the uncertainty grows and asymptotes to a level
determined by the marginal variance of the process and the predicted
value approaches the estimated marginal mean value of approximately 0
(Figure \ref{fig:R-BT-proj}). There is no uncertainty in the projected
linear trend because the values were specified as known whereas there is
a small level of uncertainty in the projections of the recent average
based on the uncertainty in the estimated mean. Comparing projected
recruitment for the northern and southern stock components, the
projected expected recruitment has much lower uncertainty than the
projected random effects regardless of whether bottom temperature
affects recruitment. For the northern component, CVs of projected
expected recruitment are smaller than those of recruitment random
effect, but there are larger differences between CVs of projected
expected recruitment assuming AR1 and the other scenarios for the future
bottom temperature anomalies are made (Figure \ref{fig:R-F-SSB-CVs}).

Because projected fishing mortality is specified to be the same as that
in the terminal year the same uncertainty is also propagated in the
projections (Figure \ref{fig:F-SSB-proj}). As we would expect, there is
no effect of alternative assumptions about bottom temperature in the
projections for the southern stock component because the effect of the
bottom temperature anomaly is only on the northern component. However,
the alternative bottom temperature assumptions do affect projections of
northern component SSB and the projections do not stabilize when the
anomalies are projected to increase linearly. When the same AR1
assumption for the bottom temperature anomalies is used in projections,
CVs of northern and total SSB projections are larger than when the
recent average or linear trend are assumed (Figure
\ref{fig:R-F-SSB-CVs}).

Unlike the annual reference points during the model years (prior to
2022), components of the \(\upphi(\widetilde{F})\) calculations are
constant in the projection years where the only changing annual values
are the recruitments used in the weighting for \(F_{40\%}\) and the
recruitments used to calculate the resulting SSB\(_{40\%}\) (Figure
\ref{fig:brps-proj}). Estimated reference points in the first few
projection years can differ depending on whether the expected
recruitment or the random effects are used in their calculation, but
longer term projected values are identical because the random effect
estimates approach the expected values as the influence of observations
diminishes. However, the estimates of uncertainty in the longer term
reference points is lower using the expected recruitment rather than the
random effects (Figures \ref{fig:annual-F40-cvs} and
\ref{fig:annual-SSB40-cvs}). The reduction in uncertainty using expected
recruitment is less for \(F_{40\%}\) than SSB\(_{40\%}\) and for the
northern component which includes bottom temperature effects than the
southern component.

\hypertarget{discussion}{%
\section*{Discussion}\label{discussion}}
\addcontentsline{toc}{section}{Discussion}

\hypertarget{black-sea-bass}{%
\subsection*{Black sea bass}\label{black-sea-bass}}
\addcontentsline{toc}{subsection}{Black sea bass}

In our investigation of bottom temperature effects on recruitment and
natural mortality and time-varying random effects on natural mortality
at age 1, we found only including effects on recruitment for the
northern stock component to be the best model with respect to AIC. Black
sea bass in the NEUS is at the northern extent of its range and the
estimated increase in expected recruitment with temperature for the
northern stock component is consistent with its range extension to the
north (McMahan et al. 2020). Opposite effects of temperature on
recruitment or population size would be expected for species in the same
general area that are at the southern extent of their range (Gabriel
1992). For example, higher recruitment of the Southern New
England-Mid-Atlantic stock of yellowtail flounder is correlated with
more cold water persistence into summer fall on the Northeast US shelf
(Sullivan et al. 2005; Miller et al. 2016b).

Our finding regarding bottom temperature effects is consistent with the
bottom temperature effects on recruitment in the assessment model
accepted during the peer-review process and which is currently used to
assess the stock (NEFSC 2023). However, AIC suggested some weight for
models with temperature effects on recruitment for the southern
component and random effects on M at age 1. Although this might suggest
making inferences on the population by using an ensemble of these and
weighting by AIC (Burnham and Anderson 2002), the estimates of
assessment outputs relevant for management (e.g., SSB and \(F\)) were
similar among the models and weighted estimates would differ very little
from the estimates with the lowest AIC. Nevertheless, continued
monitoring of effects on the southern component may be prudent.

We showed that using expected recruitment rather than recruitment random
effects to define BRPs can produce the same estimates but with lower
uncertainty. Although the option is unavailable currently in WHAM, we
expect using expected bottom temperature anomalies rather than the
random effects would also result in the same BRPs and even less
uncertainty than those we showed using expected recruitment. These
results can be important for management when uncertainty in assessment
output is used in making catch advice. The typical approach for defining
stock and harvest status is to compare BRPs defined under prevailing
conditions with current stock and harvest levels, but we could also make
the comparisons with projected BRPs like those we demonstrated are
possible. The uncertainty in these ratios could also be used in defining
catch advice.

For black sea bass, it is evident that the projection methodology of the
environmental covariate has an impact on short-term projection estimates
of recruitment and SSB (Figure 6). In the projection period, the AR1
models for bottom temperature and recuitment revert to the long-term
means of those processes. This is most likely not an appropriate
assumption for stocks in the NEUS given observations and hypothesis of
increasing water temperature and changing environmental conditions (Hare
2016; Pershing et al. 2021). Because of these concerns, the most recent
stock assessment for yellowtail flounder used a change point analysis on
environmental covariates to determine current conditions. The mean
environmental covariate from the current condition was then used as a
data input in WHAM to inform short-term projections (NEFSC, 2024).
Unfortunately, this methodology still has limitations because it is
unlikely that current environmental conditions are stable. Similarly, a
linear increase of temperature is also not likely in the future (see
past trends, Figure 6 top row).As more assessments begin including
effects of environmental covariates explicitly in assessment models,
assumptions about these covariates in calculated BRPs and population
projections becomes an important decision in the management process.

\hypertarget{general-model-aspects}{%
\subsection*{General Model aspects}\label{general-model-aspects}}
\addcontentsline{toc}{subsection}{General Model aspects}

WHAM provides a large class of models for time- and age-varying random
effects for population attributes and treatment of environmental
covariates and their effects on populations using state-space methods
(Maunder 2024). This framework accounts for the magnitude of differing
uncertainties in observations and stochastic population dynamics
processes. Version 2.0 extends these inferences to movement rates for
multi-region models and stock (component) specific configuration of
random and environmental covariate effects.

\hypertarget{future-developments}{%
\subsection*{Future developments}\label{future-developments}}
\addcontentsline{toc}{subsection}{Future developments}

An obvious limitation to WHAM is the inability to include tagging
observations of any type. Such observations are critical to estimation
of movement parameters (e.g., Goethel et al. 2019). We used estimates of
movement parameters from a companion assessment model to define prior
distributions for corresponding movement parameters in the state-space
black sea bass model, but other than the tagging data, both of the
assessment models used much the same data which can lead to
inappropriate inferences. In lieu of integrating an observation model
for the tagging data, a better approach would be to estimate the
movement parameters externally from just the tagging data.

Tagging data can also inform mortality rate parameters (Hampton 1991).
It is well known that natural mortality is seldom estimated in
assessment models because the observed data often provide little
information to distinguish natural mortality from other assessment model
parameters (Lee et al. 2011; Clark 2022). Estimation of natural
mortality may be even more challenging within state-space assessment
models with their greater flexibility from inclusion of time-varying
random effects. For example, we fixed the mean natural mortality for age
1 fish and Cadigan (2016) and Stock et al. (2021) also estimated natural
mortality deviations. However, the simulation studies of Miller et al.
(In review) suggest estimating natural mortality in WHAM is feasible in
some situations with typical assessment data. Regardless, tagging
observations can provide improved inferences and allowing tagging
observations in WHAM should be a high priority even for models with a
single stock and region. This extension would also be beneficial for
black sea bass because the current assessment uses much the same data to
estimate movement outside of the model as we are using to estimate the
other parameters in the WHAM model.

WHAM version 2.0 only allows models where each stock spawns and recruits
in a single region however, some stocks are assessed with spatial units
that are assumed to have a global stock and recruitment model that
distributes recruitment to each region (Kapur et al. 2021). Therefore,
extending WHAM to allow such models would be beneficial. Similarly age
composition data are scarce for many stocks incorporating the work by
Correa et al. (2023) on allowing length composition and modeling growth
within WHAM would provide an assessment tool to include process errors
in assessment with such data and possibly with multiple stocks and
regions.

\hypertarget{conclusions}{%
\subsection*{Conclusions}\label{conclusions}}
\addcontentsline{toc}{subsection}{Conclusions}

WHAM version 2.0 extends the existing R package to allow multiple stocks
and discrete spatial regions. It allows autocorrelated random effects
and environmental covariate effects on movement rates. Estimates of
movement rates from auxiliary studies can be used to construct prior
distributions for movement rates in lieu of integrated likelihoods for
tagging data. Version 2.0 also allows seasonal treatment of different
fishing fleets and movement of individual stocks, and internal
estimation of SPR- and MSY-based reference points accounting for
seasonal spatial dynamics including movement and mortality and any
environmental covariate effects.

We applied this extended version to investigate alternative hypotheses
about effects of bottom temperature on recruitment and age 1 natural
mortality for black sea bass. Our analyses indicate evidence for effects
of bottom temperature on recruitment for the northern component was
stronger than other models that included effects on age 1 mortality
and/or corresponding effects on the southern stock component.

Future WHAM development should prioritize including tagging data to
inform stock assessment parameter estimation and merging in the version
created by Correa et al. (2023) to allow modeling of growth and
inclusion of length and length-at-age composition observations.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-ablehales97}{}}%
Able, K.W., and Stanton Hales, Jr., L. 1997. Movements of juvenile black
sea bass \emph{{C}entropristis striata} (linnaeus) in a southern {N}ew
{J}ersey estuary. Journal of Experimental Marine Biology and Ecology
\textbf{213}(2): 153--167.
doi:\href{https://doi.org/10.1016/S0022-0981(96)02743-8}{10.1016/S0022-0981(96)02743-8}.

\leavevmode\vadjust pre{\hypertarget{ref-andersenkeiding02}{}}%
Andersen, P.K., and Keiding, N. 2002. Multi-state models for event
history analysis. Statistical Methods in Medical Research \textbf{11}:
91--115.
doi:\href{https://doi.org/10.1191/0962280202SM276ra}{10.1191/0962280202SM276ra}.

\leavevmode\vadjust pre{\hypertarget{ref-arnason72}{}}%
Arnason, A.N. 1972. Parameter estimates from mark-recapture experiments
on two populations subject to migration and death. Researches on
Population Ecology \textbf{13}: 97--113.
doi:\href{https://doi.org/10.1007/BF02521971}{10.1007/BF02521971}.

\leavevmode\vadjust pre{\hypertarget{ref-asmfc16}{}}%
ASMFC. 2016. Black sea bass stock assessment. {Available} at
http://www.asmfc.org/uploads/file/5953f11d2016BlackSeaBassStockAssmt.pdf.

\leavevmode\vadjust pre{\hypertarget{ref-belletal15}{}}%
Bell, R.J., Richardson, D.E., Hare, J.A., Lynch, P.D., and Fratantoni,
P.S. 2015. Disentangling the effects of climate, abundance, and size on
the distribution of marine fish: An example based on four stocks from
the {N}ortheast {US} shelf. ICES Journal of Marine Science
\textbf{72}(5): 1311--1322.
doi:\href{https://doi.org/10.1093/icesjms/fsu217}{10.1093/icesjms/fsu217}.

\leavevmode\vadjust pre{\hypertarget{ref-bosleyetal22}{}}%
Bosley, K.M., Schueller, A.M., Goethel, D.R., Hanselman, D.H., Fenske,
K.H., Berger, A.M., Deroba, J.J., and Langseth, B.J. 2022. Finding the
perfect mismatch: Evaluating misspecification of population structure
within spatially explicit integrated population models. Fish and
Fisheries \textbf{23}(2): 294--315.
doi:\href{https://doi.org/10.1111/faf.12616}{10.1111/faf.12616}.

\leavevmode\vadjust pre{\hypertarget{ref-brooks24}{}}%
Brooks, E.N. 2024. Pragmatic approaches to modeling recruitment in
fisheries stock assessment: A perspective. Fisheries Research
\textbf{270}: 106896.
doi:\href{https://doi.org/10.1016/j.fishres.2023.106896}{10.1016/j.fishres.2023.106896}.

\leavevmode\vadjust pre{\hypertarget{ref-burnhamanderson02}{}}%
Burnham, K.P., and Anderson, D.R. 2002. Model selection and multimodel
inference: A practical information-theoretic approach. Springer-Verlag,
New York.

\leavevmode\vadjust pre{\hypertarget{ref-cadigan16}{}}%
Cadigan, N.G. 2016. A state-space stock assessment model for northern
cod, including under-reported catches and variable natural mortality
rates. Canadian Journal of Fisheries and Aquatic Sciences
\textbf{73}(2): 296--308.
doi:\href{https://doi.org/10.1139/cjfas-2015-0047}{10.1139/cjfas-2015-0047}.

\leavevmode\vadjust pre{\hypertarget{ref-caoetal14}{}}%
Cao, J., Truesdell, S.B., and Chen, Y. 2014. Impacts of seasonal stock
mixing on the assessment of {A}tlantic cod in the {G}ulf of {M}aine.
ICES Journal of Marine Science \textbf{71}(6): 1443--1457.
doi:\href{https://doi.org/10.1093/icesjms/fsu066}{10.1093/icesjms/fsu066}.

\leavevmode\vadjust pre{\hypertarget{ref-clark22}{}}%
Clark, W.G. 2022. Why natural mortality is estimable, in theory if not
in practice, in a data-rich stock assessment. Fisheries Research
\textbf{248}: 106203.
doi:\href{https://doi.org/10.1016/j.fishres.2021.106203}{10.1016/j.fishres.2021.106203}.

\leavevmode\vadjust pre{\hypertarget{ref-commenges99}{}}%
Commenges, D. 1999. Multi-state models in epidemiology. Lifetime Data
Analysis \textbf{5}: 315--327.
doi:\href{https://doi.org/10.1023/A:1009636125294}{10.1023/A:1009636125294}.

\leavevmode\vadjust pre{\hypertarget{ref-correaetal23}{}}%
Correa, G.M., Monnahan, C.C., Sullivan, J.Y., Thorson, J.T., and Punt,
A.E. 2023. Modelling time-varying growth in state-space stock
assessments. ICES Journal of Marine Science \textbf{80}(7): 2036--2049.
doi:\href{https://doi.org/10.1093/icesjms/fsad133}{10.1093/icesjms/fsad133}.

\leavevmode\vadjust pre{\hypertarget{ref-fabrizioetal13}{}}%
Fabrizio, M.C., Manderson, J.P., and Pessutti, J.P. 2013. Habitat
associations and dispersal of black sea bass from a {M}id-{A}tlantic
{B}ight reef. Marine Ecology Progress Series \textbf{482}: 241--253.
doi:\href{https://doi.org/10.3354/meps10302}{10.3354/meps10302}.

\leavevmode\vadjust pre{\hypertarget{ref-francis14}{}}%
Francis, R.I.C.C. 2014. Replacing the multinomial in stock assessment
models: A first step. Fisheries Research \textbf{151}: 70--84.
doi:\href{https://doi.org/10.1016/j.fishres.2013.12.015}{10.1016/j.fishres.2013.12.015}.

\leavevmode\vadjust pre{\hypertarget{ref-gabriel92}{}}%
Gabriel, W.L. 1992. Persistence of demersal fish assemblages between
{C}ape {H}atteras and {N}ova {S}cotia, {N}orthwest {A}tlantic. Journal
of the Northwest Atlantic Fisheries Science \textbf{14}(1): 29--46.

\leavevmode\vadjust pre{\hypertarget{ref-goetheletal19}{}}%
Goethel, D.R., Bosley, K.M., Hanselman, D.H., Berger, A.M., Deroba,
J.J., Langseth, B.J., and Schueller, A.M. 2019. Exploring the utility of
different tag-recovery experimental designs for use in spatially
explicit, tag-integrated stock assessment models. Fisheries Research
\textbf{219}: 105320.
doi:\href{https://doi.org/10.1016/j.fishres.2019.105320}{10.1016/j.fishres.2019.105320}.

\leavevmode\vadjust pre{\hypertarget{ref-hampton91}{}}%
Hampton, J. 1991. Estimation of southern bluefin tuna \emph{{T}hunnus
maccoyii} natural mortality and movement rates from tagging experiments.
Fishery Bulletin \textbf{89}(4): 591--610.

\leavevmode\vadjust pre{\hypertarget{ref-hareetal16}{}}%
Hare, W.E.A.N., Jonathan A. AND Morrison. 2016. A vulnerability
assessment of fish and invertebrates to climate change on the northeast
u.s. Continental shelf. PLOS ONE \textbf{11}: 1--30.
doi:\href{https://doi.org/10.1371/journal.pone.0146756}{10.1371/journal.pone.0146756}.

\leavevmode\vadjust pre{\hypertarget{ref-hearnetal87}{}}%
Hearn, W.S., Sundland, R.L., and Hampton, J. 1987. Robust estimation of
the natural mortality rate in a completed tagging experiment with
variable fishing intensity. Journal Du Conseil International Pour
L'exploration De La Mer \textbf{43}: 107--117.
doi:\href{https://doi.org/10.1093/icesjms/43.2.107}{10.1093/icesjms/43.2.107}.

\leavevmode\vadjust pre{\hypertarget{ref-kapuretal21}{}}%
Kapur, M.S., Siple, M.C., Olmos, M., Privitera-Johnson, K.M., Adams, G.,
Best, J., Castillo-Jordán, C., Cronin-Fine, L., Havron, A.M., Lee, Q.,
Methot, R.D., and Punt, A.E. 2021. Equilibrium reference point
calculations for the next generation of spatial assessments. Fisheries
Research \textbf{244}: 106132.
doi:\href{https://doi.org/10.1016/j.fishres.2021.106132}{10.1016/j.fishres.2021.106132}.

\leavevmode\vadjust pre{\hypertarget{ref-kemenysnell60}{}}%
Kemeny, J.G., and Snell, J.L. 1960. Finite markov chains. D. Van
Nostrand Company, Princeton, New Jersey.

\leavevmode\vadjust pre{\hypertarget{ref-kristensenetal16}{}}%
Kristensen, K., Nielsen, A., Berg, C., Skaug, H., and Bell, B.M. 2016.
{TMB}: Automatic differentiation and {Laplace} approximation. Journal of
Statistical Software \textbf{70}: 1--21.
doi:\href{https://doi.org/10.18637/jss.v070.i05}{10.18637/jss.v070.i05}.

\leavevmode\vadjust pre{\hypertarget{ref-leeetal11}{}}%
Lee, H.-H., Maunder, M.N., Piner, K.R., and Methot, R.D. 2011.
Estimating natural mortality within a fisheries stock assessment model:
An evaluation using simulation analysis based on twelve stock
assessments. Fisheries Research \textbf{109}(1): 89--94.
doi:\href{https://doi.org/10.1016/j.fishres.2011.01.021}{10.1016/j.fishres.2011.01.021}.

\leavevmode\vadjust pre{\hypertarget{ref-legaultrestrepo99}{}}%
Legault, C.M., and Restrepo, V.R. 1999. A flexible forward
age-structured assessment program. Col. Vol. Sci. Pap. ICCAT
\textbf{49}(2): 246--253.

\leavevmode\vadjust pre{\hypertarget{ref-maunder24}{}}%
Maunder, M.N. 2024. Towards a comprehensive framework for providing
management advice from statistical inference using population dynamics
models. Ecological Modelling \textbf{498}: 110836.
doi:\href{https://doi.org/10.1016/j.ecolmodel.2024.110836}{10.1016/j.ecolmodel.2024.110836}.

\leavevmode\vadjust pre{\hypertarget{ref-mcmahanetal20}{}}%
McMahan, M.D., Sherwood, G.D., and Grabowski, J.H. 2020. Geographic
variation in life-history traits of black sea bass
(\emph{{C}entropristis striata}) during a rapid range expansion.
Frontiers in Marine Science \textbf{7}: 567758.
doi:\href{https://doi.org/10.3389/fmars.2020.567758}{10.3389/fmars.2020.567758}.

\leavevmode\vadjust pre{\hypertarget{ref-methotwetzel13}{}}%
Methot, R.D., and Wetzel, C.R. 2013. Stock {S}ynthesis: A biological and
statistical framework for fish stock assessment and fishery management.
Fisheries Research \textbf{142}(1): 86--99.
doi:\href{https://doi.org/10.1016/j.fishres.2012.10.012}{10.1016/j.fishres.2012.10.012}.

\leavevmode\vadjust pre{\hypertarget{ref-milleretal16_yoy_survival}{}}%
Miller, A.S., Shepherd, G.R., and Fratantoni, P.S. 2016a. Offshore
habitat preference of overwintering juvenile and adult black sea bass,
\emph{{C}entropristis} \emph{striata}, and the relationship to
year-class success. {PLOS} {ONE} \textbf{11}(1): e0147627.
doi:\href{https://doi.org/10.1371/journal.pone.0147627}{10.1371/journal.pone.0147627}.

\leavevmode\vadjust pre{\hypertarget{ref-millerandersen08}{}}%
Miller, T.J., and Andersen, P.K. 2008. A finite-state continuous-time
approach for inferring regional migration and mortality rates from
archival tagging and conventional tag-recovery experiments. Biometrics
\textbf{64}(4): 1196--1206.
doi:\href{https://doi.org/10.1111/j.1541-0420.2008.00996.x}{10.1111/j.1541-0420.2008.00996.x}.

\leavevmode\vadjust pre{\hypertarget{ref-milleretalinreview1}{}}%
Miller, T.J., Britten, G., Brooks, E.N., Fay, G., Hansell, A., Legault,
C.M., Li, C., Muffley, B., Stock, B.C., and Wiedenmann, J. In review.
Factors affecting reliablity of state-space age-structured assessment
models. Canadian Journal of Fisheries Science.

\leavevmode\vadjust pre{\hypertarget{ref-milleretal16}{}}%
Miller, T.J., Hare, J.A., and Alade, L.A. 2016b. A state-space approach
to incorporating environmental effects on recruitment in an
age-structured assessment model with an application to southern {New
England} yellowtail flounder. Canadian Journal of Fisheries Science
\textbf{73}(8): 1261--1270.
doi:\href{https://doi.org/10.1139/cjfas-2015-0339}{10.1139/cjfas-2015-0339}.

\leavevmode\vadjust pre{\hypertarget{ref-milleretal18}{}}%
Miller, T.J., O'Brien, L., and Fratantoni, P.S. 2018. Temporal and
environmental variation in growth and maturity and effects on management
reference points of {Georges Bank Atlantic} cod. Canadian Journal of
Fisheries Science \textbf{75}(12): 2159--2171.
doi:\href{https://doi.org/10.1139/cjfas-2017-0124}{10.1139/cjfas-2017-0124}.

\leavevmode\vadjust pre{\hypertarget{ref-millerstock20}{}}%
Miller, T.J., and Stock, B.C. 2020. The {Woods Hole Assessment Model}
({WHAM}). https://timjmiller.github.io/wham.

\leavevmode\vadjust pre{\hypertarget{ref-mohn99}{}}%
Mohn, R. 1999. The retrospective problem in sequential population
analysis: An investigation using cod fishery and simulated data. ICES
Journal of Marine Science \textbf{56}(4): 473--488.
doi:\href{https://doi.org/10.1006/jmsc.1999.0481}{10.1006/jmsc.1999.0481}.

\leavevmode\vadjust pre{\hypertarget{ref-mosershepherd09}{}}%
Moser, J., and Shepherd, G.R. 2009. Seasonal distribution and movement
of black sea bass (\emph{{C}entropristis striata}) in the {N}orthwest
{A}tlantic as determined from a mark-recapture experiment. Journal of
Northwest Atlantic Fishery Science \textbf{40}: 17--28.

\leavevmode\vadjust pre{\hypertarget{ref-musickmercer77}{}}%
Musick, J.A., and Mercer, L.P. 1977. Seasonal distribution of black sea
bass, \emph{{C}entropristis striata}, in the {M}id-{A}tlantic {B}ight
with comments on the ecology and fisheries of the species. Transactions
of the American Fisheries Society \textbf{106}(1): 12--25.
doi:\href{https://doi.org/10.1577/1548-8659(1977)106\%3C12:SDOBSB\%3E2.0.CO;2}{10.1577/1548-8659(1977)106\textless12:SDOBSB\textgreater2.0.CO;2}.

\leavevmode\vadjust pre{\hypertarget{ref-myers98}{}}%
Myers, R.A. 1998. When do environment--recruitment correlations work?
Reviews in Fish Biology and Fisheries \textbf{8}(3): 229--249.
doi:\href{https://doi.org/10.1023/A:1008828730759}{10.1023/A:1008828730759}.

\leavevmode\vadjust pre{\hypertarget{ref-nefsc22}{}}%
NEFSC. 2022a. Final report of the haddock research track assessment
working group. {Available} at
https://s3.us-east-1.amazonaws.com/nefmc.org/14b\_EGB\_Research\_Track\_Haddock\_WG\_Report\_DRAFT.pdf.

\leavevmode\vadjust pre{\hypertarget{ref-nefsc22a}{}}%
NEFSC. 2022b. Report of the american plaice research track working
group. {Available} at
https://s3.us-east-1.amazonaws.com/nefmc.org/2\_American-Plaice-WG-Report.pdf.

\leavevmode\vadjust pre{\hypertarget{ref-nefsc23}{}}%
NEFSC. 2023. Report of the black sea bass (\emph{{C}entropristis}
\emph{striata}) research track stock assessment working group.
{Available} at
https://www.mafmc.org/s/a\_2023\_BSB\_UNIT\_RTWG\_Report\_V2\_12\_2\_2023-1.pdf.

\leavevmode\vadjust pre{\hypertarget{ref-nefsc24}{}}%
NEFSC. 2024. Butterfish research track assessment report. US Dept Commer
Northeast Fish Sci Cent Ref Doc. 24-03; 191 p.

\leavevmode\vadjust pre{\hypertarget{ref-nielsenberg14}{}}%
Nielsen, A., and Berg, C.W. 2014. Estimation of time-varying selectivity
in stock assessments using state-space models. Fisheries Research
\textbf{158}: 96--101.
doi:\href{https://doi.org/10.1016/j.fishres.2014.01.014}{10.1016/j.fishres.2014.01.014}.

\leavevmode\vadjust pre{\hypertarget{ref-perreaultetal20}{}}%
Perreault, A.M.J., Wheeland, L.J., Morgan, M.J., and Cadigan, N.G. 2020.
A state-space stock assessment model for {American} plaice on the
{Grand} {Bank} of {Newfoundland}. Journal of Northwest Atlantic Fishery
Science \textbf{51}: 45--104.
doi:\href{https://doi.org/10.2960/j.v51.m727}{10.2960/j.v51.m727}.

\leavevmode\vadjust pre{\hypertarget{ref-pershingetal21}{}}%
Pershing, A.J., Alexander, M.A., Brady, D.C., Brickman, D., Curchitser,
E.N., Diamond, A.W., McClenachan, L., Mills, K.E., Nichols, O.C.,
Pendleton, D.E., Record, N.R., Scott, J.D., Staudinger, M.D., and Wang,
Y. 2021. Climate impacts on the {G}ulf of {M}aine ecosystem: A review of
observed and expected changes in 2050 from rising temperatures.
Elementa: Science of the Anthropocene \textbf{9}(1): 00076.
doi:\href{https://doi.org/10.1525/elementa.2020.00076}{10.1525/elementa.2020.00076}.

\leavevmode\vadjust pre{\hypertarget{ref-puntetal20}{}}%
Punt, A.E., Dunn, A., Elvarsson, B., Hampton, J., Hoyle, S.D., Maunder,
M.N., Methot, R.D., and Nielsen, A. 2020. Essential features of the
next-generation integrated fisheries stock assessment package: A
perspective. Fisheries Research \textbf{229}: 105617.
doi:\href{https://doi.org/10.1016/j.fishres.2020.105617}{10.1016/j.fishres.2020.105617}.

\leavevmode\vadjust pre{\hypertarget{ref-schwarzetal93}{}}%
Schwarz, C.J., Schweigert, J.F., and Arnason, A.N. 1993. Estimating
migration rates using tag-recovery data. Biometrics \textbf{49}:
177--193. doi:\href{https://doi.org/10.2307/2532612}{10.2307/2532612}.

\leavevmode\vadjust pre{\hypertarget{ref-stockmiller21}{}}%
Stock, B.C., and Miller, T.J. 2021. The {Woods Hole Assessment Model}
({WHAM}): A general state-space assessment framework that incorporates
time- and age-varying processes via random effects and links to
environmental covariates. Fisheries Research \textbf{240}: 105967.
doi:\href{https://doi.org/10.1016/j.fishres.2021.105967}{10.1016/j.fishres.2021.105967}.

\leavevmode\vadjust pre{\hypertarget{ref-stocketal21}{}}%
Stock, B.C., Xu, H., Miller, T.J., Thorson, J.T., and Nye, J.A. 2021.
Implementing two-dimensional autocorrelation in either survival or
natural mortality improves a state-space assessment model for {Southern
New England}-{Mid Atlantic} yellowtail flounder. Fisheries Research
\textbf{237}: 105873.
doi:\href{https://doi.org/10.1016/j.fishres.2021.105873}{10.1016/j.fishres.2021.105873}.

\leavevmode\vadjust pre{\hypertarget{ref-sullivanetal05}{}}%
Sullivan, M.C., Cowen, R.K., and Steves, B.P. 2005. Evidence for
atmosphere{--}ocean forcing of yellowtail flounder (\emph{{L}imanda}
\emph{ferruginea}) recruitment in the {M}iddle {A}tlantic {B}ight.
Fisheries Oceanography \textbf{14}(5): 386--399.
doi:\href{https://doi.org/10.1111/j.1365-2419.2005.00343.x}{10.1111/j.1365-2419.2005.00343.x}.

\leavevmode\vadjust pre{\hypertarget{ref-thorsonetal17}{}}%
Thorson, J.T., Johnson, K.F., Methot, R.D., and Taylor, I.G. 2017.
Model-based estimates of effective sample size in stock assessment
models using the {D}irichlet-multinomial distribution. Fisheries
Research \textbf{192}: 84--93.
doi:\href{https://doi.org/10.1016/j.fishres.2016.06.005}{10.1016/j.fishres.2016.06.005}.

\leavevmode\vadjust pre{\hypertarget{ref-thorsonetal23}{}}%
Thorson, J.T., Miller, T.J., and Stock, B.C. 2023. The
multivariate-tweedie: A self-weighting likelihood for age and length
composition data arising from hierarchical sampling designs. ICES
Journal of Marine Science \textbf{80}(10): 2630--2641.
doi:\href{https://doi.org/10.1093/icesjms/fsac159}{10.1093/icesjms/fsac159}.

\leavevmode\vadjust pre{\hypertarget{ref-varkeyetal22}{}}%
Varkey, D.A., Babyn, J., Regular, P., Ings, D.W., Kumar, R., Rogers, B.,
Champagnat, J., and Morgan, M.J. 2022. A state-space model for stock
assessment of cod (gadus morhua) stock in NAFO subdivision 3Ps. {DFO}
{Can.} {Sci.} {Advis.} {Sec.} {Res.} {Doc.} 2022/022. v + 78 p.

\leavevmode\vadjust pre{\hypertarget{ref-yingetal11}{}}%
Ying, Y., Chen, Y., Lin, L., and Gao, T. 2011. Risks of ignoring fish
population spatial structure in fisheries management. Canadian Journal
of Fisheries and Aquatic Sciences \textbf{68}(12): 2101--2120.
doi:\href{https://doi.org/10.1139/f2011-116}{10.1139/f2011-116}.

\end{CSLReferences}

\pagebreak

\pagebreak

\hypertarget{figures}{%
\section*{Figures}\label{figures}}
\addcontentsline{toc}{section}{Figures}

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{bsb_movement_diagram}

}

\caption{Diagram of intervals within the year and configuation of the dynamics of each component of the BSB population.}\label{fig:migration-diagram}
\end{figure}
\pagebreak

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{move_prior_post}

}

\caption{Prior (black) and posterior (red) distributions of movement of northern component stock from north to south (top) and south to north (bottom).}\label{fig:move-prior-posterior}
\end{figure}
\pagebreak

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{E_R_SSB_F_fig}

}

\caption{Annual observations and posterior estimates of bottom temperature anomalies (top row), annual expected and random effect recruitment estimates (second row) and SSB and average fishing mortality at ages 6 and 7 from model $M_1$. Columns define estimates regionally (spawning region for SSB) and toal SSB and fishing mortality across regions. Polygons and vertical lines represent 95\% confidence intervals.}\label{fig:E-R-SSB-F}
\end{figure}
\pagebreak

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{selectivity_re_plot}

}

\caption{Time and age-varying selectivty for fleets and indices in the northern region with autoregressive random effects.}\label{fig:selectivity-re}
\end{figure}
\pagebreak

\begin{figure}

{\centering \includegraphics[height=0.85\textheight]{brp_status_results}

}

\caption{Estimates of average equilibrium $F$ at ages 6 and 7 that produces the 40\% spawning potential ratio, equilibrium SSB at $F_{40\%}$ based on annual inputs to $\upphi(\widetilde{F})$ calculations, annual fishing and biomass status (ratios), and bivariate kobe plots of status in 2021 where reference points represent prevailing conditions with inputs to $\upphi(\widetilde{F})$ averaged over the last 5 years and recruitment is averaged over the time series. Red and blue in first 4 rows and black and gray in final row indicate alternative annual estimates of recruitment (random effect or expected). All results are based on model $M_1$. Polygons represent 95\% confidence intervals or regions.}\label{fig:annual-BRPs}
\end{figure}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_ecov_Recruit_results}

}

\caption{Annual estimates of bottom temperature in the northern region and alternative recruitment estimates (random effect or expected) by region. Estimates from years after 2021 are from projections of model $M1$ under three alternative assumptions for the bottom temperature anomalies in the northern region. Vertical dotted line is the last year of data and polygons represent 95\% confidence intervals.}\label{fig:R-BT-proj}
\end{figure}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_F_SSB}

}

\caption{Annual estimates of average fishing mortality and SSB by region and in total. Estimates in years beyond 2021 are from projecting model $M_1$ under alternative assumptions for bottom temperature anomalies in the northern region. Vertical dotted line is the last year of data and polygons represent 95\% confidence intervals.}\label{fig:F-SSB-proj}
\end{figure}

\begin{landscape}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_brps}

}

\caption{Annual estimates by region and in total of average equilibrium $F$ at ages 6 and 7 that produces the 40\% spawning potential ratio and SSB at $F_{40\%}$. Estimates in years beyond 2021 are from projecting model $M_1$ under alternative assumptions for bottom temperature anomalies in the northern region and average $\upphi(\widetilde{F})$ inputs from the last 5 years of the unprojected model. Vertical dotted line is the last year of data and polygons represent 95\% confidence intervals.}\label{fig:brps-proj}
\end{figure}

\end{landscape}

\clearpage

\setcounter{table}{0}
\renewcommand\thetable{\arabic{table}}

\hypertarget{tables}{%
\section*{Tables}\label{tables}}
\addcontentsline{toc}{section}{Tables}

\begin{table}[!h]
\centering
\caption{\label{tab:model-desc-table}Assumptions for temperature effects and random effects for age 1 natural mortality for each model.}
\centering
\begin{tabular}[t]{llll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Temperature Effect} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3}
Model & North & South & $M$ at age 1 random effects\\
\midrule
$M_{0}$ & -- & -- & none\\
$M_{1}$ & Recruitment & -- & none\\
$M_{2}$ & -- & Recruitment & none\\
$M_{3}$ & Recruitment & Recruitment & none\\
$M_{4}$ & $M$ at age 1 & -- & none\\
\addlinespace
$M_{5}$ & -- & $M$ at age 1 & none\\
$M_{6}$ & $M$ at age 1 & $M$ at age 1 & none\\
$M_{7}$ & -- & -- & time-varying\\
$M_{8}$ & Recruitment & -- & time-varying\\
$M_{9}$ & -- & Recruitment & time-varying\\
\addlinespace
$M_{10}$ & Recruitment & Recruitment & time-varying\\
$M_{11}$ & $M$ at age 1 & -- & time-varying\\
$M_{12}$ & -- & $M$ at age 1 & time-varying\\
$M_{13}$ & $M$ at age 1 & $M$ at age 1 & time-varying\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:diff-aic-table}Difference between AIC and the lowest AIC for each model by retrospective peel.}
\centering
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{8}{c}{Peel} \\
\cmidrule(l{3pt}r{3pt}){2-9}
Model & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7\\
\midrule
$M_{0}$ & 11.83 & 11.39 & 10.41 & 10.05 & 9.86 & 9.41 & 8.47 & 8.22\\
$M_{1}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
$M_{2}$ & 12.63 & 12.03 & 11.41 & 11.06 & 10.91 & 10.34 & 9.20 & 9.22\\
$M_{3}$ & 0.80 & 0.64 & 1.00 & 1.01 & 1.05 & 0.93 & 0.73 & 1.00\\
$M_{4}$ & 13.81 & 13.35 & 12.26 & 11.82 & 11.71 & 11.18 & 10.08 & 9.85\\
\addlinespace
$M_{5}$ & 13.25 & 12.55 & 11.68 & 11.21 & 11.07 & 10.41 & 9.86 & 9.71\\
$M_{6}$ & 15.22 & 14.51 & 13.52 & 12.97 & 12.91 & 12.17 & 11.46 & 11.33\\
$M_{7}$ & 14.32 & 13.75 & 12.17 & 11.30 & 11.43 & 10.55 & 9.71 & 9.42\\
$M_{8}$ & 2.25 & 2.10 & 1.47 & 0.96 & 1.29 & 0.91 & 0.98 & 0.99\\
$M_{9}$ & 15.12 & 14.39 & 13.17 & 12.31 & 12.48 & 11.49 & 10.44 & 10.43\\
\addlinespace
$M_{10}$ & 3.05 & 2.74 & 2.47 & 1.97 & 2.34 & 1.84 & 1.71 & 1.99\\
$M_{11}$ & 16.29 & 15.74 & 14.17 & 13.30 & 13.43 & 12.55 & 11.68 & 11.39\\
$M_{12}$ & 15.73 & 14.91 & 13.43 & 12.46 & 12.64 & 11.55 & 11.10 & 10.92\\
$M_{13}$ & 17.70 & 16.91 & 15.43 & 14.46 & 14.64 & 13.55 & 13.07 & 12.88\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:rho-table}Mohn's $\rho$ for SSB, and average F at ages 6 and 7 in northern and southern regions.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{SSB} & \multicolumn{2}{c}{Average F} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Model & North & South & North & South\\
\midrule
$M_{0}$ & -0.040 & -0.023 & 0.041 & -0.048\\
$M_{1}$ & -0.040 & -0.023 & 0.041 & -0.048\\
$M_{2}$ & -0.040 & -0.024 & 0.041 & -0.047\\
$M_{3}$ & -0.040 & -0.024 & 0.041 & -0.047\\
$M_{4}$ & -0.041 & -0.022 & 0.042 & -0.048\\
\addlinespace
$M_{5}$ & -0.041 & -0.023 & 0.042 & -0.048\\
$M_{6}$ & -0.042 & -0.023 & 0.042 & -0.048\\
$M_{7}$ & -0.042 & -0.022 & 0.043 & -0.048\\
$M_{8}$ & -0.044 & -0.023 & 0.044 & -0.048\\
$M_{9}$ & -0.042 & -0.024 & 0.043 & -0.048\\
\addlinespace
$M_{10}$ & -0.044 & -0.024 & 0.044 & -0.047\\
$M_{11}$ & -0.041 & -0.022 & 0.042 & -0.048\\
$M_{12}$ & -0.042 & -0.023 & 0.043 & -0.048\\
$M_{13}$ & -0.041 & -0.023 & 0.042 & -0.048\\
\bottomrule
\end{tabular}
\end{table}
\begin{table}

\caption{\label{tab:beta-sig-peel-table}Estimates of temperature effects on recruitment and variance and autocorrelation parameters for recruitment for northern ($N$) and southern ($S$) components for models with no effects ($M_0$), effects on specific components ($M_1$ and $M_2$) or both components simultaneously ($M_3$) from the full model (Peel 0) and each retrospective peel.}
\centering
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{8}{c}{Peel} \\
\cmidrule(l{3pt}r{3pt}){2-9}
Parameter & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7\\
\midrule
$M_{1}$ $\widehat{\beta}_{R,N}$ & 0.474 & 0.480 & 0.485 & 0.476 & 0.464 & 0.468 & 0.439 & 0.445\\
$M_{2}$ $\widehat{\beta}_{R,S}$ & 0.099 & 0.105 & 0.094 & 0.095 & 0.092 & 0.096 & 0.099 & 0.091\\
$M_{3}$ $\widehat{\beta}_{R,N}$ & 0.474 & 0.480 & 0.485 & 0.476 & 0.464 & 0.468 & 0.439 & 0.445\\
$M_{3}$ $\widehat{\beta}_{R,S}$ & 0.099 & 0.105 & 0.094 & 0.095 & 0.092 & 0.096 & 0.099 & 0.091\\
$M_{0}$ Conditional $\widehat{\sigma}_{R,N}$ & 0.925 & 0.953 & 0.978 & 0.988 & 0.978 & 1.010 & 0.981 & 1.007\\
\addlinespace
$M_{1}$ Conditional $\widehat{\sigma}_{R,N}$ & 0.730 & 0.752 & 0.780 & 0.786 & 0.775 & 0.801 & 0.785 & 0.800\\
$M_{0}$ $\widehat{\rho}_{R,N}$ & 0.362 & 0.375 & 0.385 & 0.405 & 0.424 & 0.432 & 0.427 & 0.431\\
$M_{1}$ $\widehat{\rho}_{R,N}$ & 0.296 & 0.307 & 0.334 & 0.373 & 0.394 & 0.406 & 0.428 & 0.429\\
$M_{0}$ Marginal $\widehat{\sigma}_{R,N}$ & 0.992 & 1.028 & 1.060 & 1.080 & 1.081 & 1.120 & 1.084 & 1.117\\
$M_{1}$ Marginal $\widehat{\sigma}_{R,N}$ & 0.764 & 0.791 & 0.827 & 0.848 & 0.843 & 0.877 & 0.869 & 0.885\\
\bottomrule
\end{tabular}
\end{table}

\clearpage

\hypertarget{appendix-a}{%
\section*{Appendix A}\label{appendix-a}}
\addcontentsline{toc}{section}{Appendix A}

\setcounter{table}{0}
\renewcommand\thetable{A\arabic{table}}
\input{symboltable.tex}

\pagebreak

\hypertarget{supplemental-materials}{%
\section*{Supplemental Materials}\label{supplemental-materials}}
\addcontentsline{toc}{section}{Supplemental Materials}

\hypertarget{deriving-the-prior-distribution-for-movement-parameters}{%
\subsection*{Deriving the prior distribution for movement
parameters}\label{deriving-the-prior-distribution-for-movement-parameters}}
\addcontentsline{toc}{subsection}{Deriving the prior distribution for
movement parameters}

The Working Group fit a Stock Synthesis model (Methot and Wetzel 2013)
that included tagging data with 2 seasons (6 months each) and 2 regions
where a proportion \(\mu^*_1\) of the northern component moves to the
south in one season and some proportion \(\mu^*_{2\rightarrow 1}\) move
back to the south in the second season (NEFSC 2023). The seasonal
movement matrices for each season are \begin{equation*}
\boldsymbol{\mu}^*_{1} = 
  \begin{bmatrix}
     1-\mu^*_{1\rightarrow 2} & \mu^*_{1\rightarrow 2} \\
     0 & 1 \\
  \end{bmatrix}
\end{equation*} and \begin{equation*}
\boldsymbol{\mu}_{2} = 
  \begin{bmatrix}
     1 &  0 \\
     \mu^*_{2\rightarrow 1} & 1-\mu^*_{2\rightarrow 1} \\
  \end{bmatrix}.
\end{equation*}

To obtain estimates of movement proportions for the monthly intervals in
the WHAM model, the half-year movement matrices were converted to
monthly movement matrices by taking the root \(z_k\) of
\(\boldsymbol{\mu}^*_{k}\) which are defined by the number of months of
movement for each season (5 and 4, respectively). The roots of the
matrices are calculated using an eigen decomposition of the matrices
\[ \boldsymbol{\mu}_k =  \left(\boldsymbol{\mu}_k^*\right)^{z_k} = \mathbf{V}_k \mathbf{D}_k^{z_k} \mathbf{V}_k^{-1}\]
where \(z_1 = 1/5\) for and \(z_2 = 1/4\), and \(\mathbf{V}_{k}\) and
\(\mathbf{D}_{k}\) are the matrix of eigenvectors (columnwise) and the
diagonal matrix of corresponding eigenvalues of
\(\boldsymbol{\mu}^*_k\). The Working Group used a parametric bootstrap
approach to determine an appropriate standard deviation for the prior
distribution for the movement parameters. Stock Synthesis also estimates
parameters on a transformed scale, but different from WHAM:
\[\mu^*_{r\rightarrow r'} = \frac{1}{1 + 2e^{-x_{r\rightarrow r'}}}\]
The estimated parameters and standard errors from the Stock Synthesis
model were \(x_{1\rightarrow 2}=-1.44\) and \(x_{2\rightarrow 1}=1.94\)
and \(SE(x_{1\rightarrow 2})) = 0.21\) and
\(SE(x_{2\rightarrow 1})) = 0.37\). The resulting in the estimated
proportions were \(\mu^*_{1\rightarrow 2}=0.11\) and
\(\mu^*_{2\rightarrow 1}=0.78\).

In WHAM, an additive logit transformation is used which is simply a
logit transformation when there are only two regions: \[
\mu_{r\rightarrow r'} = \frac{1}{1+e^{-y_{r\rightarrow r'}}}.
\] We simulated 1000 values from a normal distribution with mean and
standard deviation defined by the parameter estimate and standard error
\(\tilde x_{{r\rightarrow r'},b} \sim N(x_{r\rightarrow r'}, SE(x_{r\rightarrow r'}))\)
from the Stock Synthesis model. For each simulated value we constructed
\(\tilde {\boldsymbol{\mu}}^*_{{r\rightarrow r'},b}\), took the
appropriate root and calculated inverse logit for
\(\tilde y_{{r\rightarrow r'},b}\). We calculated the mean and standard
deviation of the values \(y_{i,b}\). The mean values did not differ
meaningfully from the transformation of the original estimates
(\(y_{1\rightarrow 2} = -3.79\) and \(y_{2\rightarrow 1} = -0.79\)) and
the standard deviation was approximately 0.2 for both parameters.

\hypertarget{diagnostics}{%
\subsection*{Diagnostics}\label{diagnostics}}
\addcontentsline{toc}{subsection}{Diagnostics}

\hypertarget{jitter-fits-for-model-m_0}{%
\subsubsection*{\texorpdfstring{Jitter fits for model
\(M_0\)}{Jitter fits for model M\_0}}\label{jitter-fits-for-model-m_0}}
\addcontentsline{toc}{subsubsection}{Jitter fits for model \(M_0\)}

WHAM by default completes three newton steps after the stats::nlminb
minimization function completes to reduce the gradient at the minimized
NLL. However, this generally has negligible effects on model estimates
and the NLL. To reduce computation time, we did not complete these
newton steps when performing jitter fits of the model. Without the
Newton steps, the maximum (absolute) gradient sizes are generally less
than 0.01 for models that converge satisfactorily.

The 50 jitter fits demonstrated that a local minimum was obtained for
the original fit of model \(M_0\) (Figure \ref{fig:jitter-M0}). Some
lower NLLs were obtained with unacceptable gradients, but a slightly
lower NLL was found with a satisfactory gradient and with and invertible
hessian. We therefore refit model \(M_0\) and all remaining models using
the better parameter estimates as initial values.

\hypertarget{jitter-fits-for-model-m_1}{%
\subsubsection*{\texorpdfstring{Jitter fits for model
\(M_1\)}{Jitter fits for model M\_1}}\label{jitter-fits-for-model-m_1}}
\addcontentsline{toc}{subsubsection}{Jitter fits for model \(M_1\)}

The 50 jitter fits gave no evidence of a better minimization of the NLL.
Three lower NLLs were obtained, but with unacceptably large gradients
(Figure \ref{fig:jitter-M1}). The largest differences in parameter
estimates for these three jitters were for numbers at age and
selectivity random effects variance and correlation parameters.

\hypertarget{self-test-for-model-m_1}{%
\subsubsection*{\texorpdfstring{Self test for model
\(M_1\)}{Self test for model M\_1}}\label{self-test-for-model-m_1}}
\addcontentsline{toc}{subsubsection}{Self test for model \(M_1\)}

Initial fits to simulated data from model \(M_1\) showed estimation of
the observation error standard deviation multiplier for the recreational
catch-per angler indices in the north and south regions was unstable.
Many of the fits to the simulated data produced implausible estimates at
the 0 boundary for these parameters (very negative values on log-scale).
Therefore, we completed self-tests with these parameters fixed at the
true values.

For 7 of the the simulated data sets the model failed to optimize. The
maximum absolute gradient was \(<10^{-6}\) for only 9 and \(<10^{-4}\)
for 52 of the 93 successfully fitted models. The poor convergence
appeared to be attributable to the estimation of the scalar for the
standard errors of the log-transformed northern Recreational CPA index
for which estimates tended to 0 for nearly all of the fits (\(<0.01\)
for 83 fits). However, even across all fits including those with poor
convergence, the SSB estimates appeared to be reliable (Figure
\ref{fig:self-test-fig}).

\setcounter{table}{0}
\renewcommand\thetable{S\arabic{table}}

\begin{landscape}\begin{table}

\caption{\label{tab:age-comp-sel-table}Configuration of age composition likelihoods, mean selectivity models, and selectivity random effects models for each age composition data component. For all logistic-normal likelihoods, any ages observed as zeros are treated as missing.}
\centering
\begin{tabular}[t]{llll}
\toprule
Data component & Age Composition Likelihood & Mean Selectivity model & Random effects Model\\
\midrule
North commercial fleet & Dirichlet-Multinomial & age-specific (ages > 3 fully selected) & AR1 correlation by age and year\\
North recreational fleet & Logistic-normal (Independent) & age-specific (ages > 6 fully selected) & AR1 correlation by age and year\\
South commercial fleet & Logistic-normal (AR1 correlation) & logistic & None\\
South recreational fleet & Logistic-normal (AR1 correlation) & logistic & None\\
North recreational CPA index & Logistic-normal (Independent) & age-specific (ages > 1 fully selected) & AR1 correlation by year\\
\addlinespace
North VAST index & Dirichlet-Multinomial & age-specific (ages > 4 fully selected) & AR1 correlation by age and year\\
South recreational CPA index & Logistic-normal (AR1 correlation) & age-specific (ages > 2 fully selected) & None\\
South VAST index & Logistic-normal (AR1 correlation) & age-specific (ages > 1 fully selected) & None\\
\bottomrule
\end{tabular}
\end{table}
\end{landscape}

\begin{table}

\caption{\label{tab:aic-wts-table}Model AIC weights for each retrospective peel.}
\centering
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{8}{c}{Peel} \\
\cmidrule(l{3pt}r{3pt}){2-9}
Model & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7\\
\midrule
$M_{0}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.01 & 0.01\\
$M_{1}$ & 0.45 & 0.43 & 0.42 & 0.38 & 0.41 & 0.37 & 0.36 & 0.38\\
$M_{2}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
$M_{3}$ & 0.30 & 0.31 & 0.25 & 0.23 & 0.24 & 0.23 & 0.25 & 0.23\\
$M_{4}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
\addlinespace
$M_{5}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
$M_{6}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
$M_{7}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
$M_{8}$ & 0.15 & 0.15 & 0.20 & 0.24 & 0.21 & 0.24 & 0.22 & 0.23\\
$M_{9}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
\addlinespace
$M_{10}$ & 0.10 & 0.11 & 0.12 & 0.14 & 0.13 & 0.15 & 0.15 & 0.14\\
$M_{11}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
$M_{12}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
$M_{13}$ & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00\\
\bottomrule
\end{tabular}
\end{table}

\setcounter{figure}{0}
\renewcommand\thefigure{S\arabic{figure}}

\pagebreak

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{fit_0_jitter_plt}

}

\caption{Minimized negative log-likelihood for 50 fits where minimization used initial parameter values jittered from those provided by an initial fit for model $M_0$. Black jitters had maximum absolute gradient values < $10^{-10}$, grey jitters had values > $10^{-10}$ and < 1, and red jitters had values > 1.}\label{fig:jitter-M0}
\end{figure}
\pagebreak

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{fit_1_jitter_plt}

}

\caption{Minimized negative log-likelihood for 50 fits where minimization used initial parameter values jittered from those provided by an initial fit for model $M_1$. Fits with black dots had maximum absolute gradient value < 0.01 and fits with red dots had values > 10.}\label{fig:jitter-M1}
\end{figure}
\pagebreak

\begin{landscape}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{self_test_results}

}

\caption{Median relative bias of SSB for the North and South stock components for estimation models where the observation variance of log-indices are fixed and estimation is by maximum marginal likelihood (ML) or Restricted Maximum Likelihood (REML) or where the standard deviation of logistic-normal age composition observations are assumed to be 0.1 (over sqrt(1000)) (Low Error). Black and Red dashed lines represent the median of the yeaerly medians and the median across all yearly relative errors, respectively.}\label{fig:self-test-fig}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{BTA_full_fig}

}

\caption{Observations with 95\% confidence intervals (points with vertical lines) and posterior estimates with 95\% confidence intervals (lines with polygons) of bottom temperature anomalies in the north and south regions from model $M_1$. Gray vertical line defines the first year that the black sea bass stock is modeled.}\label{fig:bottom-temperature}
\end{figure}

\end{landscape}
\pagebreak

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{best_R_Ecov}

}

\caption{Expected and random effect recruitment estimates for the northern stock component. Color of points defined by the corresponding annual bottom temperature anomaly.}\label{fig:BT-Ecov-R}
\end{figure}
\pagebreak

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{SSB_F_R_rel_M1}

}

\caption{Estimates of SSB, F, and recruitment relative to those of the best performing model, $M_1$.}\label{fig:SSB-F-R-rel-M1}
\end{figure}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{Ecov_M1_rel_M0}

}

\caption{Relative differences in posterior estimates of northern region bottom temperature anomalies ($\widehat X$) from the null model without effects on recruitment ($M_0$) and with effects on the northern stock component ($M_1$).}\label{fig:Ecov-M1-rel-M0}
\end{figure}

\begin{landscape}

\begin{figure}

{\centering \includegraphics[height=0.9\textheight]{R_SSB_F_cv_results}

}

\caption{Coefficients of variation for estimates of alternative recruitment estimates (random effects or expected), average fishing mortality at age 6 and 7, and SSB by region and in total from model $M_1$. Values in years after 2021 are from projecting model $M_1$ under three alternative assumptions for the bottom temperature anomolies.}\label{fig:R-F-SSB-CVs}
\end{figure}
\end{landscape}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_F40_CV}

}

\caption{Coefficients of variation for annual equilibrium average $F$ at ages 6 and 7 that produces the 40\% spawning potential ratio as a function of annual expected recruitment or recruitment random effects and annual inputs to $\upphi(\widetilde{F})$ calculations. Estimates in years after 2021 are from projecting model $M_1$ under three alternative assumptions for the bottom temperature anomolies.}\label{fig:annual-F40-cvs}
\end{figure}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_SSB40_CV}

}

\caption{Coefficients of variation for annual equilibrium SSB$_{40\%}$ as a function of annual expected recruitment or recruitment random effects and annual inputs to $\upphi(\widetilde{F})$ calculations and alternative annual recruitment types. Estimates in years after 2021 are from projecting model $M_1$ under three alternative assumptions for the bottom temperature anomolies.}\label{fig:annual-SSB40-cvs}
\end{figure}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_F_status_results}

}

\caption{Annual estimates of ratios of fishing mortality to $F_{40\%}$ by region and in total. Estimates in years beyond 2021 are from projecting model $M_1$ under alternative assumptions for bottom temperature anomalies in the northern region. Vertical dotted line is the last year of data and polygons represent 95\% confidence intervals.}\label{fig:F-status-proj}
\end{figure}
\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_SSB_status_results}

}

\caption{Annual estimates of ratios of SSB to SSB$_{40\%}$ by region and in total. Estimates in years beyond 2021 are from projecting model $M_1$ under alternative assumptions for bottom temperature anomalies in the northern region. Vertical dotted line is the last year of data and polygons represent 95\% confidence intervals.}\label{fig:SSB-status-proj}
\end{figure}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_SSB_status_CV}

}

\caption{Coefficients of variation for annual ratios of SSB and equilibrium SSB$_{40\%}$ where the latter is a function of annual expected recruitment or recruitment random effects and annual inputs to $\upphi(\widetilde{F})$ calculations. Estimates in years after 2021 are from projecting model $M_1$ under three alternative assumptions for the bottom temperature anomolies.}\label{fig:annual-SSB-status-cvs}
\end{figure}

\begin{figure}

{\centering \includegraphics[height=0.95\textheight]{proj_F_status_CV}

}

\caption{Coefficients of variation for annual ratios of average fishing mortality and equilibrium $\bar{F}_{40\%}$ at ages 6 and 7 where the latter is a function of annual expected recruitment or recruitment random effects and annual inputs to $\upphi(\widetilde{F})$ calculations. Estimates in years after 2021 are from projecting model $M_1$ under three alternative assumptions for the bottom temperature anomolies.}\label{fig:annual-F-status-cvs}
\end{figure}

\end{document}